---
title: "Analysis1"
output: html_document
date: "2026-03-28"
---

```{r setup, message=FALSE, warning=FALSE}
library(naniar)
library(tidyverse)
library(dplyr)
library(gridExtra)
library(corrplot)
library(pheatmap)
library(cluster)
library(factoextra)
library(Hmisc)
library(mice)
library(here)
here::here()

freshdata <- read_csv(
  here("data","processed","freshdata_cleaned1.csv"),
  name_repair = "minimal"
)
```

```{r}
#minor data changes - dropping, deriving, relocating parameters
freshdata <- freshdata %>%
  select(-Baseline_follicle_size, -Second_scan_follicle_size, -Final_follicle_size)
```

## Filter first cycle and binary encoding 
```{r}
freshdata <- freshdata %>%
  filter(Cycle_no == 1)

print(table(freshdata$Treatment_type))


## DELETE THIS AT THE END
write_csv(
  freshdata,
  here("data", "processed", "freshdata_cleaned_new.csv")
)
```

## Outcome distribution
```{r}
ggplot(freshdata, aes(x = No_mature_eggs)) +
  geom_histogram(binwidth = 2, fill = "light blue", colour = "black") +
  theme_minimal() +
  labs(title = "Distribution of Mature Oocytes",
       x = "Number of Mature Oocytes",
       y = "Count")


ggplot(freshdata, aes(x = Total_stim_dose)) +
  geom_histogram(binwidth = 200, fill = "steelblue", colour = "black") +
  theme_minimal() +
  labs(title = "Distribution of Total Stimulation Dose",
       x = "Total Stim Dose (iu)",
       y = "Count")
```

```{r}
summary(freshdata$No_mature_eggs)

sd(freshdata$No_mature_eggs, na.rm = TRUE) #8.1419
var(freshdata$No_mature_eggs, na.rm = TRUE) #66.29054
mean(freshdata$No_mature_eggs, na.rm = TRUE) #11.24359
```

## Distributions
```{r}
plot_histograms <- function(df, bins = 20, exclude_binary = TRUE) {
  library(ggplot2)
  library(dplyr)
  #Select numeric columns
  #num_df <- df %>% select(where(is.numeric))
  #Loop through columns
  for (col in names(df)) {
    clean_label <- col %>%
      str_replace_all("_", " ") %>%
      str_to_sentence()
    p <- ggplot(df, aes(x = .data[[col]])) +
      geom_histogram(bins = bins, fill = "steelblue", color = "black") +
      labs(
        title = paste("Distribution of", col),
        x = col,
        y = "Count"
      ) +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black")
      )
    print(p)
  }
}

plot_histograms(num_vars)
```

## Missingness
```{r}
#Table: percentage missing per variable, sorted descending
miss_table <- freshdata %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "Parameter", values_to = "Pct_missing") %>%
  arrange(desc(Pct_missing)) %>%
  filter(Pct_missing > 0)
 
print(miss_table, n = 50)
```

```{r}
ggplot(miss_table, aes(x = reorder(Parameter, Pct_missing), y = Pct_missing)) +
  geom_col(fill = "light blue") +
  coord_flip() +
  labs(title = "Percentage missing per variable",
       x = NULL, y = "% missing") +
  theme_minimal()
```
```{r, fig.width=12, fig.height=7}
freshdata |>
  select(where(~ any(is.na(.)))) |>
  vis_miss(sort_miss = TRUE, cluster = TRUE) +
  labs(
    title = "Missingness Heatmap",
    subtitle = "Variables sorted by missingness, rows clustered by pattern of missingness",
    x = "Parameters",
    y = "Observations"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1, size = 10),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10, colour = "Black"),
    legend.position = "bottom"
  )
```
## Correlation
```{r}
num_vars <- freshdata %>% 
  select(
    Age, 
    BMI,
    Baseline_AMH,
    Baseline_total_follicles,
    Baseline_left_follicles,
    Baseline_right_follicles,
    Baseline_endometrium,
    GnRH_antagonist_start_day,
    GnRH_antagonist_end_day,
    GnRH_antagonist_duration,
    GnRH_antagonist_total_dose,
    FSH_total_dose,
    FSH_duration,
    FSH_initial_dose,
    FSH_final_dose,
    FSH_max_dose,
    FSH_min_dose,
    FSH_dose_changes,
    FSH_change_day_1,
    FSH_change_day_2,
    #FSH_change_day_3,
    FSH_dose_delta,
    FSH_avg_daily,
    hMG_total_dose,
    hMG_duration,
    hMG_initial_dose,
    hMG_final_dose,
    hMG_max_dose,
    hMG_min_dose,
    hMG_dose_changes,
    hMG_change_day_1, 
    #hMG_change_day_2,
    #hMG_change_day_3,
    hMG_dose_delta,
    hMG_avg_daily,
    hCG_trigger_dose,
    GnRH_agonist_trigger_dose,
    Last_day_of_stim,
    Final_total_follicles,
    Final_left_follicles,
    Final_right_follicles,
    Final_follicles_lessthan_11.9,
    Final_endometrium_ec,
    Day_of_ec,
    Pre_wash_sperm_conc,
    Pre_wash_sperm_motility,
    Pre_wash_normal_sperm_morphology,
    Post_wash_sperm_conc,
    Post_wash_sperm_motility,
    Semen_vol,
    No_eggs,
    No_mature_eggs,
    No_fertilised,
    Fertilisation_rate
  )
```

## Bin age ranges
```{r}
freshdata <- freshdata %>%
  mutate(
    Age_group = case_when(
      Age < 30 ~ "<30",
      Age >= 30 & Age < 35 ~ "30–34",
      Age >= 35 & Age < 38 ~ "35–37",
      Age >= 38 & Age < 40 ~ "38–39",
      Age >= 40 ~ "40+"
    )
  )
```

```{r, fig.width=13, fig.height=8}
cor_matrix1 <- cor(num_vars, method = "spearman", use = "pairwise.complete.obs")
corrplot(cor_matrix1,
         method = "color",
         type = "upper",
         #order = "hclust",
         tl.col = "black",
         tl.cex = 0.6,
         col = colorRampPalette(c("blue", "white", "red"))(200))
```

```{r, fig.width=12, fig.height=7}
res1 <- rcorr(as.matrix(num_vars), type = "spearman")

corrplot(res1$r,
         method = "color",
         type = "upper",
         tl.col = "black",
         tl.cex = 0.6,
         col = colorRampPalette(c("blue", "white", "red"))(200),
         p.mat = res1$P,
         sig.level = 0.05,
         insig = "blank")
```

```{r}
cor_mat1 <- res1$r
p_mat1 <- res1$P

#tidy up 
cor_df1 <- as.data.frame(as.table(cor_mat1)) %>%
  rename(Var1 = Var1, Var2 = Var2, Correlation = Freq) %>%
  mutate(p_value = as.vector(p_mat1)) %>%
  filter(Var1 != Var2)

#remove duplicates
cor_df1 <- cor_df1 %>%
  rowwise() %>%
  mutate(pair = paste(sort(c(Var1, Var2)), collapse = "_")) %>%
  ungroup() %>%
  distinct(pair, .keep_all = TRUE)
```

## T1: most sig correlations by p-value
```{r}
top_p1 <- cor_df1 %>%
  arrange(p_value) %>%
  slice(1:10)
```

## T2: strongest correlations by maginitude of r
```{r}
top_r1 <- cor_df1 %>%
  arrange(desc(abs(Correlation))) %>%
  slice(1:10)
```

## T3: top sig
```{r}
top_r_sig1 <- cor_df1 %>%
  filter(p_value < 0.05) %>%
  arrange(desc(abs(Correlation))) %>%
  slice(1:10)
```

## Correlations with number of mature oocytes
```{r}
cor_with_outcome1 <- data.frame(
  Variable = colnames(res1$r),
  Correlation = res1$r["No_mature_eggs", ],
  p_value = res1$P["No_mature_eggs", ]
)

cor_with_outcome1 <- cor_with_outcome1 %>%
  filter(Variable != "No_mature_eggs")

cor_with_outcome1 %>%
  arrange(desc(abs(Correlation)))
```
```{r}
cor_with_outcome1 %>%
  arrange(p_value)
```

## smaller correlation matrices
```{r}
baseline_vars <- c(
    "Unique_number",
    "Aetiology_group",
    "Age", 
    "BMI",
    "Baseline_AMH",
    "Baseline_total_follicles",
    "Baseline_endometrium",
    "Protocol_type",
    "No_mature_eggs"
    )

baseline_df <- freshdata %>%
  select(all_of(baseline_vars))

protocol_vars <- c(
  "GnRH_antagonist_start_day",
  "GnRH_antagonist_end_day",
  "GnRH_antagonist_duration",
  "GnRH_antagonist_total_dose",
  "FSH_total_dose",
  "FSH_duration",
  "FSH_initial_dose",
  "FSH_final_dose",
  "FSH_max_dose",
  "FSH_min_dose",
  "FSH_dose_changes",
  "FSH_change_day_1",
  "FSH_change_day_2",
    #FSH_change_day_3,
  "FSH_dose_delta",
  "FSH_avg_daily",
  "hMG_total_dose",
  "hMG_duration",
  "hMG_initial_dose",
  "hMG_final_dose",
  "hMG_max_dose",
  "hMG_min_dose",
  "hMG_dose_changes",
  "hMG_change_day_1", 
    #hMG_change_day_2,
    #hMG_change_day_3,
  "hMG_dose_delta",
  "hMG_avg_daily",
  "hCG_trigger_dose",
  "GnRH_agonist_trigger_dose",
  "Last_day_of_stim",
  "No_mature_eggs"
)

protocol_df <- freshdata %>%
  select(all_of(protocol_vars))

male_vars <- c(
  "Pre_wash_sperm_conc",
  "Pre_wash_sperm_motility",
  "Pre_wash_normal_sperm_morphology",
  "Post_wash_sperm_conc",
  "Post_wash_sperm_motility",
  "Semen_vol",
  "No_mature_eggs",
  "No_fertilised",
  "Fertilisation_rate"
)

male_df <- freshdata %>%
  select(all_of(male_vars))
```

```{r}
cor_matrix2 <- cor(baseline_df, method = "spearman", use = "pairwise.complete.obs")
corrplot(cor_matrix2,
         method = "color",
         type = "upper",
         #order = "hclust",
         tl.col = "black",
         tl.cex = 0.8,
         col = colorRampPalette(c("blue", "white", "red"))(200))
```
```{r}
res2 <- rcorr(as.matrix(baseline_df), type = "spearman")
cor_with_outcome2 <- data.frame(
  Variable = colnames(res2$r),
  Correlation = res2$r["No_mature_eggs", ],
  p_value = res2$P["No_mature_eggs", ]
)

cor_with_outcome2 <- cor_with_outcome2 %>%
  mutate(abs_r = abs(Correlation)) %>%
  arrange(desc(abs_r))

cor_with_outcome2 %>%
  select(Variable, Correlation, p_value)
```

```{r, fig.width=12, fig.height=8}
cor_matrix3 <- cor(protocol_df, method = "spearman", use = "pairwise.complete.obs")
corrplot(cor_matrix3,
         method = "color",
         type = "upper",
         #order = "hclust",
         tl.col = "black",
         tl.cex = 0.7,
         col = colorRampPalette(c("blue", "white", "red"))(200))
```

```{r}
cor_matrix4 <- cor(male_df, method = "spearman", use = "pairwise.complete.obs")
corrplot(cor_matrix4,
         method = "color",
         type = "upper",
         #order = "hclust",
         tl.col = "black",
         tl.cex = 0.8,
         col = colorRampPalette(c("blue", "white", "red"))(200))
```


```{r}
ggplot(freshdata, aes(x = Age, y = No_mature_eggs)) +
  
  geom_point(alpha = 1, size = 1.5, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between Age and Number of Mature Oocytes",
    x = "Age (years)",
    y = "Number of Mature Oocytes"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.4),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Baseline_AMH, y = No_mature_eggs)) +
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between AMH Levels and Number of Mature Oocytes",
    x = "AMH (Anti Mullerian Hormone) Levels",
    y = "Number of Mature Oocytes"
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r, fig.width=10, fig.height=6}
ggplot(freshdata, aes(x = Baseline_total_follicles, y = No_mature_eggs)) +
  
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  # Add trend line (very important)
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between Baseline Total Follicles (prior to first IVF cycle) and Number of Mature Oocytes",
    x = "Age (years)",
    y = "Number of Total Baseline Follicles"
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Age, y = Baseline_AMH)) +
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between Age and AMH Levels",
    x = "Age",
    y = "AMH Levels"
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Age, y = No_mature_eggs, colour = Protocol_type)) +
  geom_point(alpha = 0.9, size = 2) +
    scale_colour_manual(values = c(
    "#D55E00", "#56B4E9", "#CC79B9", "#F0E442"
    )) +
  #geom_smooth(method = "lm", se = FALSE) +
  #scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Age vs Mature Oocytes by Protocol Type",
    x = "Age",
    y = "Number of Mature Oocytes",
    colour = "Protocol"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black")
  )
```


```{r}
ggplot(freshdata, aes(x = Age, y = No_mature_eggs, colour = Aetiology_group)) +
  geom_point(alpha = 0.9, size = 2) +
  scale_colour_manual(values = c(
    "#D55E00", "#56B4E9", "#CC79B9", "#F0E442"
    #"#0072B2", "#D55E00", "#CC79A7", "#000000"
  )) +
  labs(
    title = "Age vs Number of Mature Oocytes by Aetiology",
    x = "Age (years)",
    y = "Number of Mature Oocytes",
    colour = "Aetiology"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.title = element_text(face = "bold")
  )
```

```{r}
ggplot(freshdata, aes(x = FSH_total_dose, y = No_mature_eggs, colour = Aetiology_group)) +
  geom_point(alpha = 0.9, size = 2) +
  scale_colour_manual(values = c(
  "#D55E00", "#56B4E9", "#CC79B9", "#F0E442"
  #"#0072B2", "#D55E00", "#CC79A7", "#000000"
  )) +
  #Trend lines per protocol
  #geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Total FSH dose vs Number of Mature Oocytes by Aetiology",
    x = "Total FSH dose (iu)",
    y = "Number of Mature Oocytes",
    colour = "Aetiology"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.title = element_text(face = "bold")
  )
```
```{r}
ggplot(freshdata, aes(x = Total_stim_dose, y = No_mature_eggs, colour = Aetiology_group)) +
  geom_point(alpha = 0.9, size = 2) +
  scale_colour_manual(values = c(
  "#D55E00", "#56B4E9", "#CC79B9", "#F0E442"
  #"#0072B2", "#D55E00", "#CC79A7", "#000000"
  )) +
  #Trend lines per protocol
  #geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Total Stimulation dose vs Number of Mature Oocytes by Aetiology",
    x = "Total Stimulation dose (iu)",
    y = "Number of Mature Oocytes",
    colour = "Aetiology"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.title = element_text(face = "bold")
  )
```

```{r}
ggplot(freshdata, aes(x = FSH_total_dose, y = No_mature_eggs)) +
  geom_point(alpha = 0.5, colour = "#1F78B4") +
  #geom_smooth(method = "loess", se = FALSE, colour = "darkred") +
  facet_wrap(~ Age_group) +
  labs(
    title = "FSH Dose vs Mature Oocytes Across Age Groups",
    x = "Total FSH Dose",
    y = "Number of Mature Oocytes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    strip.text = element_text(face = "bold"),
    axis.text = element_text(color = "black")
  )


ggplot(freshdata, aes(x = FSH_total_dose, y = No_mature_eggs, colour = Age_group)) +
  
  geom_point(alpha = 0.9, size = 2) +
  
  #geom_smooth(method = "lm", se = FALSE) +
  
  scale_colour_brewer(palette = "Dark2") +
  
  labs(
    title = "FSH Dose vs Mature Oocytes by Age Group",
    x = "Total FSH Dose",
    y = "Number of Mature Oocytes",
    colour = "Age Group"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black")
  )
```

```{r}
ggplot(freshdata, aes(x = hMG_total_dose, y = No_mature_eggs, colour = Age_group)) +
  
  geom_point(alpha = 0.9, size = 2) +
  
  #geom_smooth(method = "lm", se = FALSE) +
  
  scale_colour_brewer(palette = "Dark2") +
  
  labs(
    title = "hMG Dose vs Mature Oocytes by Age Group",
    x = "Total hMG Dose (iu)",
    y = "Number of Mature Oocytes",
    colour = "Age Group"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black")
  )
```



```{r}
ggplot(freshdata, aes(x = Baseline_AMH, y = FSH_total_dose)) +
  
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between AMH levels and FSH dose administered",
    x = "Baseline AMH levels",
    y = "Total FSH dose (iu)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Baseline_AMH, y = hMG_total_dose)) +
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between AMH levels and hMG dose administered",
    x = "Baseline AMH levels",
    y = "Total hMG dose (iu)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Age, y = FSH_total_dose)) +
  
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between Age and FSH dose administered",
    x = "Age (years)",
    y = "Total FSH dose (iu)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Age, y = hMG_total_dose)) +
  
  geom_point(alpha = 1, size = 2, colour = "#1F78B4") +
  #geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between Age and Total hMG dose administered",
    x = "Age (years)",
    y = "Total FSH dose (iu)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Protocol_type, y = No_mature_eggs)) +
  geom_boxplot(fill = "#A6CEE3", colour = "black", outlier.alpha = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1.5) +
  labs(
    title = "Number of Mature Oocytes by Protocol Type",
    x = "Protocol Type",
    y = "Number of Mature Oocytes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

```{r}
ggplot(freshdata, aes(x = Aetiology, y = No_mature_eggs)) +
  geom_boxplot(fill = "#A6CEE3", colour = "black", outlier.alpha = 0.5) +
  # Add jittered points (VERY useful for your dataset size)
  geom_jitter(width = 0.2, alpha = 0.4, size = 1.5) +
  labs(
    title = "Number of Mature Oocytes by Protocol Type",
    x = "Protocol Type",
    y = "Number of Mature Oocytes"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
```

### MODEL
```{r}
library(MASS)

model1 <- glm.nb(
  No_mature_eggs ~ Age + FSH_total_dose + hMG_total_dose +
    Baseline_AMH + BMI,
  data = freshdata
)

summary(model1)
exp(coef(model1))
```


```{r}
model2 <- glm.nb(No_mature_eggs ~ Baseline_AMH * FSH_total_dose + Age, data = freshdata)
summary(model2)
```
```{r}
model3 <- glm.nb(No_mature_eggs ~ Age * FSH_total_dose + Baseline_AMH, data = freshdata)
summary(model3)
```

##Model dosing
```{r}
model4 <- glm.nb(FSH_avg_daily ~ Baseline_AMH + Age, data = freshdata)
summary(model4)
```
## Model dose strategy not just dose

```{r}
model5 <- glm.nb(
  No_mature_eggs ~ Baseline_AMH + BMI + FSH_avg_daily + FSH_dose_changes + hMG_total_dose + hMG_duration,
  data = freshdata
)
summary(model5)
```
## Model testing whether age affects AMH levels

```{r}
lm1 <- lm(log(Baseline_AMH) ~ Age, data = freshdata)
summary(lm1)
```
```{r}
plot(lm1, which = 2)
plot(lm1, which = 3)
plot(lm1, which = 4)  # Cook’s distance
```

```{r}
freshdata$Aetiology_group <- factor(freshdata$Aetiology_group)

lm2 <- lm(log(Baseline_AMH) ~ Age * Aetiology_group, data = freshdata)

plot(lm2, which = 2)
plot(lm2, which = 3)
plot(lm2, which = 4)

summary(lm2)
```

```{r}
ggplot(freshdata, aes(Age, log(Baseline_AMH + 1), colour = Aetiology_group)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Age",
    y = "log(AMH + 1)",
    colour = "Aetiology"
  ) +
  theme_minimal()

freshdata <- freshdata %>%
  filter(!is.na(Aetiology_group))

ggplot(freshdata, aes(Age, log(Baseline_AMH), colour = Aetiology_group)) +
  geom_point(alpha = 0.75, size = 1.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title = "Effect of Age on AMH Across Aetiology Groups",
    x = "Age (years)",
    y = "log(AMH)",
    colour = "Aetiology group"
  ) +
  theme_classic(base_size = 14)
```

```{r}
library(MASS)
nb1 <- glm.nb(No_mature_eggs ~ Age * Aetiology_group, data = freshdata)
summary(nb1)
exp(coef(nb1))
exp(confint(nb1))
```

```{r}
nb2 <- glm.nb(No_mature_eggs ~ Age + log(Baseline_AMH) + Aetiology_group, data = freshdata)
summary(nb2)
plot(nb2)
```

```{r}
nb3 <- glm.nb(No_mature_eggs ~ log(Baseline_AMH) + Age * Aetiology_group, data = freshdata)
summary(nb3)
```

```{r}
nb4 <- glm.nb(No_mature_eggs ~ Age + log(Baseline_AMH) + Aetiology_group + FSH_avg_daily + hMG_avg_daily + GnRH_antagonist_duration, data = freshdata)
summary(nb4)
```

## Modelling Dose


```{r}
lm3 <- lm(FSH_total_dose ~ Baseline_AMH + Age + Aetiology_group, data = freshdata)
summary(lm3)

plot(lm3)
```

```{r}
lm4 <- lm(FSH_avg_daily ~ Age, data = freshdata)
summary(lm4)
```
```{r}
freshdata$Treatment_type <- factor(freshdata$Treatment_type)

nb5 <- glm.nb(No_mature_eggs ~ Total_stim_dose + Treatment_type + Age, data = freshdata)
summary(nb5)
```

```{r}
lm5 <- lm(log(Total_stim_dose) ~ Age * Aetiology_group + Treatment_type, data = freshdata)
summary(lm5)
plot(lm5)


hist(log(freshdata$Total_stim_dose))
```

```{r}
nb6 <- glm.nb(No_mature_eggs ~ Baseline_AMH * Total_stim_dose + Age + Aetiology_group, data = freshdata)
summary(nb6)
##higher dose does not mean more eggs
```


```{r}
freshdata <- freshdata %>%
  mutate(
    AMH_group = cut(
      Baseline_AMH,
      breaks = quantile(Baseline_AMH, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE),
      labels = c("Low AMH", "Medium AMH", "High AMH"),
      include.lowest = TRUE
    )
  )

ggplot(freshdata, aes(Total_stim_dose, No_mature_eggs, colour = AMH_group)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title = "Effect of Stimulation Dose on Mature Oocytes by AMH Level",
    x = "Total Stimulation Dose (IU)",
    y = "Number of Mature Oocytes",
    colour = "AMH group"
  ) +
  theme_classic(base_size = 14)
```




