---
title: "Cleaning pipeline"
output: html_document
date: "2026-03-12"
---

```{r setup, message=FALSE, warning=FALSE}

#Load required packages
library(tidyverse)
library(dplyr)
library(stringr)
library(janitor)

library(here)
here::here()

#Load functions scripts needed for this pipeline
source(here("scripts", "cleaning_helpers.R"))
source(here("scripts", "cleaning_functions_parsers.R"))
```

```{r}
#list.files("..")
#list.files("../scripts")
#list.files("../data/raw")
```

```{r, message = FALSE}
freshdata <- read_csv(
  here("data","raw","patientdata_fresh1.csv"),
  name_repair = "minimal"
)

#freshdata <- read.csv("../data/raw/patientdata_fresh1.csv")
#df 521x269
```

```{r}
#only 90 have data currently
#removing rows 91-176
freshdata[90, ]
freshdata[91, ]
freshdata[521, ]
freshdata <- freshdata[-c(91:521), ]
```

```{r}
freshdata <- freshdata %>%
  clean_names() %>%
  rename_with(~ str_to_title(.))
```

```{r}
freshdata[, 61:ncol(freshdata)]
freshdata <- freshdata %>%
  select(1:60)

#Define columns to drop - can be modified depending on what is needed
drop_cols <- c(
  "Consented_date",
  "Date_fluid_obtained",
  "Ethnicity",
  "Dob",
  "Baseline_scan",
  "Days_of_stim",
  "Day_of_2nd_scan",
  "Date_of_additional_scan",
  "Dx_last_scan_prior_to_trigger",
  "Day_of_trigger",
  "Date_of_ec",
  "Icsi_ivf"
)

freshdata <- freshdata %>%
  select(
    -any_of(drop_cols)
  )

names(freshdata)
#90x39 df
```

## Rename parameters (colnames)

```{r}
##new name = old name
freshdata <- freshdata %>%
  rename(
    Male_factor = Male_factor_y_n,
    Funding = Nhs_pp,
    Infertility_duration = Duration_of_infertility_months,
    Age = Age_at_present_cycle,
    Baseline_FSH = Baseline_fsh, 
    Baseline_AMH = Baseline_amh,
    Baseline_AFC = Baseline_afc, 
    Baseline_E2 = Baseline_e2,
    Baseline_endometrium = Baseline_endometrium_mm,
    Baseline_left_follicles = No_of_follicles_on_bl_scan_left,
    Baseline_right_follicles = No_of_follicles_on_bl_scam_right,
    Baseline_follicle_size = Size_of_follicle,
    Second_scan_left_follicles = X2nd_scan_no_of_follciles_left,
    Second_scan_right_follicles = X2nd_scan_no_of_follciles_right,
    Second_scan_follicle_size = X2nd_scan_follicle_size,
    Second_scan_endometrium = X2nd_endo_mm,
    Final_left_follicles = Follicles_left,
    Final_right_follicles = Follicles_right,
    Final_follicle_size = Foll_size_mm,
    Final_endometrium_ec = Endo_at_ec_mm,
    Pre_wash_sperm_conc = Pre_wash_sperm_concentration_mil_ml,
    Pre_wash_sperm_motility = Pre_wash_sperm_motility_percent,
    Pre_wash_normal_sperm_morphology = Pre_wash_sperm_morphology_normal_forms_percent,
    Post_wash_sperm_conc = Post_wash_sperm_concentration_mil_ml,
    Post_wash_sperm_motility = Post_wash_sperm_motility_percent,
    Semen_vol = Semen_volume_ml,
    Sperm_prep = Sperm_prep_dg_mini_none,
    Sperm_mobility = Sperm_mobil_y_n, 
    Sperm_source = Sperm_source_f_fz
  )
```

```{r}
freshdata <- standardise_missing_values(freshdata)
```

## Derive PGD

```{r}
freshdata <- freshdata %>%
  mutate(
    ##Derive PGD column
    PGD = if_else(
      str_detect(Aetiology, regex("\\bPGD\\b", ignore_case = TRUE)),
      "Y",
      "N"
    ),
    .after = Aetiology
  )
```

## Aetiology

```{r}
freshdata <- freshdata %>%
  mutate(
    Aetiology = clean_aetiology(Aetiology)
  )
```

## Aetiology QC Checks

```{r}
print(unique(freshdata$Aetiology))

##Check = 0
freshdata %>%
  filter(
    str_detect(Aetiology, "PGD") &
    PGD != "Y"
  )

freshdata %>%
  filter(
    PGD == "Y" &
    Male_factor == "Y" &
    Aetiology == "No Female Factor"
  )

#0
freshdata %>%
  filter(str_detect(Aetiology, "PGD"))
```

## Funding

```{r}
unique(freshdata$Funding)

##Clean
freshdata <- freshdata %>%
  mutate(
    Funding = str_squish(Funding),   # remove extra spaces
    Funding = case_when(
      Funding %in% c("", "missing") ~ NA_character_,
      Funding == "PP (self funded)" ~ "PP",
      TRUE ~ Funding
    )
  )

##QC checks
unique(freshdata$Funding)
freshdata %>% 
  count(Funding)
```

## Male Factor
```{r}
unique(freshdata$Male_factor)

freshdata <- freshdata %>%
  mutate(
    Male_factor = str_to_title(Male_factor)
  )

#QC
unique(freshdata$Male_factor)
```

## Infertility Duration

```{r}
unique(freshdata$Infertility_duration)

##check > entries = 6
sum(grepl("^>", freshdata$Infertility_duration), na.rm = TRUE)

##clean column
freshdata <- freshdata %>%
  mutate(
    Infertility_duration = str_remove(Infertility_duration, "^>"),
    Infertility_duration = as.numeric(Infertility_duration)
  )

## QC checks
freshdata %>% 
  count(Infertility_duration)
```

## Treatment

```{r}
print(unique(freshdata[ ,"Treatment"]))

freshdata <- freshdata %>%
  mutate(
    Treatment = case_when(
      Treatment %in% c("IVF/ICSI", "ICSI/IVF") ~ "IVF/ICSI",
      Treatment %in% c("ICSI-38", "ICSI (Surrogacy)") ~ "ICSI",
      Treatment %in% c("IVF (35)", "IVF short with pills") ~ "IVF",
      Treatment == "" ~ NA_character_,
      TRUE ~ Treatment
    )
  )

print(unique(freshdata[ ,"Treatment"]))
```

```{r}
str(freshdata$Age) #num
```

## Baselines

```{r}
print(unique(freshdata$Baseline_FSH))

freshdata <- freshdata %>%
  mutate(
    across(
      c(Baseline_FSH, Baseline_AMH, Baseline_AFC, Baseline_E2),
      clean_baseline
    ),
    Baseline_endometrium = round(as.numeric(Baseline_endometrium), 2),
    across(
      c(Baseline_right_follicles, Baseline_left_follicles),
      ~ as.numeric(.x)
    )
  )
```

## Deriving total follicle count

```{r}
freshdata <- freshdata %>%
  mutate(
    Baseline_total_follicles = Baseline_right_follicles + Baseline_left_follicles
  ) %>%
  relocate(
    Baseline_total_follicles,
    .after = Baseline_E2
  )
```

## Protocol

```{r}
unique(freshdata$Protocol)

freshdata <- freshdata %>%
  mutate(
    Protocol = Protocol %>%
      str_squish() %>%
      str_replace_all("(\\d+)\\s+iu", "\\1iu") %>%      # 225 iu to 225iu
      str_replace("hMG\\+rRFSH", "hMG+rFSH") %>%        # typo fix
      str_replace("Fertility preservation", "Fertility Preservation")
  )

##Derive 3 columns for clarity
freshdata <- freshdata %>%
  mutate(
    ##1.Protocol type
    Protocol_type = case_when(
      str_detect(Protocol, "Long Agonist") ~ "Long Agonist",
      str_detect(Protocol, "Short") ~ "Short Antagonist",
      str_detect(Protocol, "Egg Donor") ~ "Egg Donor",
      str_detect(Protocol, "Fertility Preservation") ~ "Fertility Preservation",
      TRUE ~ NA_character_
    ),
    ##2.Drug
    Drug = case_when(
      str_detect(Protocol, "hMG\\+rFSH") ~ "Dual",
      str_detect(Protocol, "rFSH") & !str_detect(Protocol, "hMG") ~ "rFSH",
      str_detect(Protocol, "hMG") & !str_detect(Protocol, "rFSH") ~ "hMG",
      TRUE ~ NA_character_
    ),
    ##3.Dose
    Dose_iu = as.numeric(str_extract(Protocol, "\\d+"))
    
  ) %>%
  relocate(Protocol_type, Drug, Dose_iu, .after = Protocol)


##Checks
freshdata %>% count(Protocol_type) #59 Short, 6 Long, 4 Fertility Preservation, 21 egg donor
freshdata %>% count(Drug) #30 rFSH, 15 hMG, 20 Dual, 25 NA

##Drop original Protocol column, and derived Drug and Dose_iu columns as information is redundant and repeats later
freshdata <- freshdata %>%
  select(-Protocol, -Dose_iu, -Drug)
```

## Cycle
```{r}
str(freshdata$Cycle_no)
```

## Second scans

```{r}
freshdata <- freshdata %>%
  mutate(
    Second_scan_endometrium = round(as.numeric(Second_scan_endometrium), 2),
    across(
      c(Second_scan_left_follicles, Second_scan_right_follicles),
      ~ as.numeric(.x)
    )
  )
```

## FSH and HmG

```{r}
#Clean
freshdata <- freshdata %>%
  mutate(
    across(c(Fsh_iu, H_mg), clean_stim_string)
  )

#check
print(unique(freshdata$Fsh_iu))

#manual corrections
freshdata <- freshdata %>%
  mutate(
    Fsh_iu = case_when(
      Unique_number == "8029938C" ~ "225 (D1-9)",
      Unique_number == "8239838A" ~ "225 (D1-5), 150 (D6-8), 225 (D9-12)",
      TRUE ~ Fsh_iu
    )
  )
print(unique(freshdata$Fsh_iu)) #30 different FSH doses

print(unique(freshdata$H_mg)) #16 different H_mg doses

##Parse and derive parameters - FSH
freshdata <- freshdata %>%
  mutate(parsed = map(Fsh_iu, parse_stimulation, prefix = "FSH")) %>%
  unnest(parsed)
#relocate
freshdata <- freshdata %>%
  relocate(starts_with("FSH_"), .after = Fsh_iu)

##Parse and derive parameters - hMG
freshdata <- freshdata %>%
  mutate(parsed = map(H_mg, parse_stimulation, prefix = "hMG")) %>%
  unnest(parsed)
#relocate
freshdata <- freshdata %>%
  relocate(starts_with("hMG_"), .after = H_mg)

#drop original Fsh_iu and H_mg columns
freshdata <- freshdata %>%
  select(-Fsh_iu, -H_mg, -Total_hmg, -Total_gonal_f_iu)
```

## Trigger

```{r}
freshdata <- freshdata %>%
  mutate(Trigger_h_cg = clean_trigger_string(Trigger_h_cg))

##Derive parameters from triggers
freshdata <- freshdata %>%
  mutate(
    #Extract drugs --> trigger mechanism
    has_hcg = str_detect(Trigger_h_cg, regex("Ovitrelle", ignore_case = TRUE)),
    has_agonist = str_detect(Trigger_h_cg, regex("Buserelin", ignore_case = TRUE)),
    #Extract dose
    hCG_trigger_dose = ifelse(
      has_hcg,
      as.numeric(str_extract(
        str_extract(Trigger_h_cg, "Ovitrelle\\s*\\d+"),
      "\\d+"
      )),
      0
    ),
    GnRH_agonist_trigger_dose = ifelse(
      has_agonist,
      as.numeric(str_extract(
        str_extract(Trigger_h_cg, "Buserelin\\s*\\d+"),
        "\\d+"
      )),
      0
    ),
    #Mechanism
    Trigger_mechanism = case_when(
      has_hcg & has_agonist ~ "Dual",
      has_hcg ~ "hCG",
      has_agonist ~ "GnRH_agonist",
      TRUE ~ NA_character_
    )
  ) %>%
  select(-has_hcg, -has_agonist)

##Relocate
freshdata <- freshdata %>%
  relocate(
    Trigger_mechanism,
    hCG_trigger_dose,
    GnRH_agonist_trigger_dose,
    .after = Trigger_h_cg
  )

#Remove original H_cg_trigger column
freshdata <- freshdata %>%
  select(-Trigger_h_cg)
```

## Other medication

```{r}
print(unique(freshdata$Other_medications))

##9 patients with other medications entered, info entered in wrong way, should be dual trigger sometimes
##use to manually correct and to change parameters and derive parameters
othermeds <- freshdata %>%
  filter(!is.na(Other_medications) & Other_medications != "") %>%
  select(
    Unique_number,
    Other_medications,
    Total_other_iu,
    Trigger_mechanism,
    GnRH_agonist_trigger_dose,
    hCG_trigger_dose,
    Last_day_of_stim
  )

freshdata <- freshdata %>%
  mutate(
    Trigger_mechanism = case_when(
      Unique_number %in% c(
        "2653224K",
        "8119398Y",
        "8120863D",
        "8308873N",
        "8356317B",
        "8391436J"
      ) ~ "Dual",
      TRUE ~ Trigger_mechanism
    ),
    
    GnRH_agonist_trigger_dose = case_when(
      Unique_number == "2653224K" ~ 700,
      Unique_number == "8119398Y" ~ 32,
      Unique_number == "8120863D" ~ 240,
      Unique_number == "8308873N" ~ 240,
      Unique_number == "8356317B" ~ 260,
      Unique_number == "8391436J" ~ 240,
      TRUE ~ GnRH_agonist_trigger_dose
    )
  )
```

## Derive progesterone

```{r}
#progesterone used --> yes or no 
freshdata <- freshdata %>% 
  mutate( Progesterone = ifelse(str_detect(Other_medications, "Provera"), 1, 0) )

freshdata <- freshdata %>% 
  mutate(Progesterone = ifelse(is.na(Progesterone), 0, Progesterone)) 

freshdata <- freshdata %>% 
  mutate( Progesterone_total_dose = ifelse(Progesterone == 1, Total_other_iu, 0) )

str(freshdata$Progesterone)
str(freshdata$Progesterone_total_dose)

freshdata <- freshdata %>%
  mutate(
    Progesterone_total_dose = as.numeric(as.character(Progesterone_total_dose)),
  )
```

## Fyramadel --> GnRH antagonist

```{r}
#Clean
freshdata <- freshdata %>%
  mutate(
    across(c(Fyremadel), clean_stim_string)
  )

print(unique(freshdata$Fyremadel)) #8 unique doses 

#Parse and derive parameters
freshdata <- freshdata %>%
  mutate(parsed_fyr = map(Fyremadel, parse_fyr)) %>%
  unnest(parsed_fyr)
#relocate
freshdata <- freshdata %>%
  relocate(starts_with("Fyr_"), .after = Fyremadel)

##check
freshdata %>%
  select(Fyremadel, Fyr_used, Fyr_start_day, Fyr_end_day, Fyr_duration, Fyr_total_dose) %>%
  slice_sample(n = 30)

#Restructuring Fyremadel --> GnRH antagonist, allows for different medication names
freshdata <- freshdata %>%
  rename_with(~ gsub("^Fyr_", "GnRH_antagonist_", .x), starts_with("Fyr_"))

freshdata <- freshdata %>%
  select(-Fyremadel, -Total_fyremadel_iu)
```

## More other medications

```{r}
#in the other medications column there are other GnRH antagonist drugs: Cetrotide 
#manual editing by unique number
freshdata <- freshdata %>%
  mutate(
    GnRH_antagonist_used = if_else(
      Unique_number == "8228365W",
      1,
      GnRH_antagonist_used
    ),
    GnRH_antagonist_start_day = if_else(
      Unique_number == "8228365W",
      6,
      GnRH_antagonist_start_day
    ),
    GnRH_antagonist_end_day = if_else(
      Unique_number == "8228365W",
      13,
      GnRH_antagonist_end_day
    ),
    GnRH_antagonist_duration = if_else(
      Unique_number == "8228365W",
      8,
      GnRH_antagonist_duration
    ),
    GnRH_antagonist_total_dose = if_else(
      Unique_number == "8228365W",
      2000,
      GnRH_antagonist_total_dose
    )
  )
```

## Fixing introduced NAs due to derived parameters

```{r}
#Replace with 0
freshdata <- freshdata %>%
  mutate(across(
    matches("FSH_|Hmg_|GnRH_antagonist"),
    ~replace_na(., 0)
  ))

freshdata <- freshdata %>%
  mutate(across(
    starts_with("Hmg_change_day"),
    as.integer
  ))
```

## Final scans
```{r}
freshdata <- freshdata %>%
  mutate(
    Final_endometrium_ec = round(as.numeric(Final_endometrium_ec), 2),
    across(
      c(Final_left_follicles, Final_right_follicles),
      ~ as.numeric(.x)
    )
  )
```

## Sperm parameters

```{r}
sperm_cols <- c("Pre_wash_sperm_conc", "Pre_wash_sperm_motility", "Pre_wash_normal_sperm_morphology", "Post_wash_sperm_conc", "Post_wash_sperm_motility", "Semen_vol")

freshdata <- freshdata %>%
  mutate(
    across(all_of(sperm_cols), clean_sperm_numeric)
  )

#Manual correction
freshdata <- freshdata %>%
  mutate(
    Pre_wash_sperm_motility = ifelse(
      Pre_wash_sperm_motility == "Only 2 twitching seen on Makler",
      0,
      Pre_wash_sperm_motility
    )
  )

#reorganize
freshdata <- freshdata %>%
  relocate(
    Sperm_source,
    .after = Day_of_ec
  )

#sperm source
print(unique(freshdata$Sperm_source))

##can later convert to binary
freshdata <- freshdata %>%
  mutate(
    Sperm_source = case_when(
      stringr::str_detect(Sperm_source, "FZ \\(Donor\\)") ~ "FZ_donor",
      stringr::str_detect(Sperm_source, "FZ \\(Partner\\)") ~ "FZ_partner",
      Sperm_source == "F" ~ "F",
      TRUE ~ NA_character_
    )
  )

print(unique(freshdata$Sperm_prep))
```

## Days

```{r}
freshdata <- freshdata %>%
  mutate(
    across(c(Last_day_of_stim, Day_of_ec), parse_day)
  )
```

```{r}
freshdata <- freshdata %>%
  mutate(
    No_fertilised = as.numeric(as.character(No_fertilised)),
  )
```

## Derive fertilisation rate
```{r}
freshdata <- freshdata %>%
  mutate(
    Fertilisation_rate = No_fertilised / No_mature_eggs
  )
```

## Drop redundant parameter and reorganise

```{r}
freshdata <- freshdata %>%
  select(-Other_medications, -Total_other_iu, -Trigger_mechanism)

freshdata <- freshdata %>%
  relocate(
    FSH_used,
    hMG_used,
    .after = Protocol_type
  )

freshdata <- freshdata %>%
  relocate(
    Progesterone,
    Progesterone_total_dose,
    .after = hMG_avg_daily
  )

str(freshdata)
cat("Final dataset size:", dim(freshdata), "\n")
```

## Save

```{r}
write_csv(
  freshdata,
  here("data", "processed", "freshdata_cleaned.csv")
)

names(freshdata)
```
