---
title: "Cleaning + Merging with new data"
output: html_document
date: "2026-03-23"
---

```{r setup, message=FALSE, warning=FALSE}
#Load required packages
library(tidyverse)
library(dplyr)
library(stringr)
library(janitor)

library(here)
here::here()

#Load functions scripts needed for this pipeline
source(here("scripts", "cleaning_helpers.R"))
source(here("scripts", "cleaning_functions_parsers.R"))
```

```{r setup, message=FALSE, warning=FALSE}
update_data <- read_csv(
  here("data","raw","Final_Fresh_Merged.csv"),
  name_repair = "minimal"
)

update_data1 <- read_csv(
  here("data","raw","Missing_Fresh_Data.csv"),
  name_repair = "minimal"
)

smalldata <- read_csv(
  here("data", "processed", "freshdata_cleaned1.csv"),
  name_repair = "minimal"
)
```

## Standardise missing values
```{r}
update_data <- standardise_missing_values(update_data)
update_data1 <- standardise_missing_values(update_data1)

#freshdata <- standardise_missing_values(freshdata)
```

## MERGE
```{r}
##difference in data structure of this parameter so will fail to merge
str(update_data$FRZ_D7_EevaScores)
str(update_data1$FRZ_D7_EevaScores)

#ensure both are numeric
update_data1 <- update_data1 %>%
  mutate(
    FRZ_D7_EevaScores = trimws(FRZ_D7_EevaScores),
    FRZ_D7_EevaScores = na_if(FRZ_D7_EevaScores, ""),
    FRZ_D7_EevaScores = as.numeric(FRZ_D7_EevaScores)
  )

#Merge 
data <- bind_rows(update_data, update_data1)
```

```{r}
data <- data %>%
  clean_names() %>%
  rename_with(~ str_to_title(.))
```

## Renaming parameters of interest  
```{r}
data <- data %>%
  rename(
    Unique_number = Unique_nbr,
    Funding = Patient_class,
    Age = Patient_age_at_td1,
    BMI = Bmi,
    Aetiology = Causeof_infertility,
    Protocol_type = Protocol,
    Final_total_follicles = Total_fol_on_last_scan,
    Final_follicles_lessthan_11.9 = Total_fol_11_9mm_on_last_scan,
    Baseline_AMH = Anti_mullerian_hormone_result,
    Pre_wash_sperm_motility = Initial_motile,
    Pre_wash_sperm_conc = Initial_con,
    Post_wash_sperm_conc = Post_con,
    Post_wash_sperm_motility = Post_motile,
    Semen_vol = Volume,
    Sperm_prep = Spz_prep,
    Treatment = Art_method,
    Fertilisation_rate = Fert_rate,
    No_eggs = Nbr_retrieved,
    No_inseminated = Nbr_inseminated,
    No_mature_eggs = Nbr_miid0,
    No_mature_eggs_16hr = Nbr_mii16hr,
    No_fertilised_eggs = Nbr_2pn
  )

#filter by cycle start date --> check that they are dates
#infertility_duration, cycle start date, total FSH dose, total hMG dose, trigger med, total trigger dose, fert method, 
```

# Extract parameters of interest
```{r}
parameters <- c(
    "Unique_number",
    "Funding",
    "Age",
    "BMI",
    "Infertility_duration",
    "Aetiology",
    "Cycle_start_date",
    "Protocol_type",
    "Total_fsh_dose",
    "Totalh_mg_dose",
    "Trigger_med",
    "Total_trigger_dose",
    "Final_total_follicles",
    "Final_follicles_lessthan_11.9",
    "Natural_follicle_info",
    "Baseline_AMH",
    "Treatment",
    "Pre_wash_sperm_conc",
    "Pre_wash_sperm_conc",
    "Post_wash_sperm_motility",
    "Post_wash_sperm_conc",
    "Semen_vol",
    "Sperm_prep",
    "Pelvic_date",
    "No_eggs",
    "No_inseminated",
    "No_mature_eggs",
    "No_mature_eggs_16hr",
    "No_fertilised_eggs",
    "Fertilisation_rate"
  )

freshdata <- data %>%
  dplyr::select(
    all_of(parameters),
    72:107 #follicle size counts too --> to derive baseline total follicle count, potentially
    )
```

## Funding
```{r}
print(unique(freshdata$Funding))

freshdata <- freshdata %>%
  mutate(Funding = clean_funding(Funding))
```

## Age and BMI
```{r}
str(freshdata$Age)
str(freshdata$BMI)
str(freshdata$Infertility_duration)
```

## Derive PGD
```{r}
freshdata <- freshdata %>%
  mutate(
    ##Derive PGD column
    PGD = if_else(
      str_detect(Aetiology, regex("\\bPGD\\b", ignore_case = TRUE)),
      "Y",
      "N"
    ),
    .after = Aetiology
  )
```

## Derive Male Factor
```{r}
freshdata <- freshdata %>%
  mutate(
    ##Derive Male factor column
    Male_factor = if_else(
      !is.na(Aetiology) & str_detect(Aetiology, regex("\\bMale Factor\\b", ignore_case = TRUE)),
      "Y",
      "N"
    ),
    .after = Aetiology
  )
```

## Aetiology
```{r}
print(unique(freshdata$Aetiology))
```
```{r}
freshdata <- freshdata %>%
  mutate(
    Aetiology = clean_aetiology1(Aetiology)
  )

print(unique(freshdata$Aetiology))
```
## Protocol

```{r}
print(unique(freshdata$Protocol_type))

freshdata <- freshdata %>%
  mutate(
    protocol_lower = stringr::str_to_lower(stringr::str_squish(Protocol_type)),
    #Cleaning function
    protocol_type_clean = clean_protocol_type(protocol_lower),
    
    ##if the drug is mentioned in the protocol use 1
    FSH_used = ifelse(
      stringr::str_detect(protocol_lower, "fsh|rfsh") |
      (!is.na(Total_fsh_dose) & Total_fsh_dose > 0),
      1, 0
    ),
    
    hMG_used = ifelse(
      stringr::str_detect(protocol_lower, "hmg") |
      (!is.na(Totalh_mg_dose) & Totalh_mg_dose > 0),
      1, 0
    ),
    
    FSH_starting_dose = extract_fsh_dose(protocol_lower),
    hMG_starting_dose = extract_hmg_dose(protocol_lower),
    
    .after = Protocol_type
  ) %>%
  select(-protocol_lower)

##Fixing 0s and NAs
freshdata <- freshdata %>%
  mutate(
    #FSH starting dose
    FSH_starting_dose = case_when(
      FSH_used == 0 ~ 0,
      FSH_used == 1 & !is.na(FSH_starting_dose) ~ FSH_starting_dose,
      FSH_used == 1 & is.na(FSH_starting_dose) ~ NA_real_
    ),
    #hMG starting dose
    hMG_starting_dose = case_when(
      hMG_used == 0 ~ 0,
      hMG_used == 1 & !is.na(hMG_starting_dose) ~ hMG_starting_dose,
      hMG_used == 1 & is.na(hMG_starting_dose) ~ NA_real_
    )
  )

freshdata <- freshdata %>%
  mutate(
    Total_fsh_dose = case_when(
      FSH_used == 0 ~ 0,
      FSH_used == 1 & !is.na(Total_fsh_dose) ~ Total_fsh_dose,
      FSH_used == 1 & is.na(Total_fsh_dose) ~ NA_real_
    ),
    
    Totalh_mg_dose = case_when(
      hMG_used == 0 ~ 0,
      hMG_used == 1 & !is.na(Totalh_mg_dose) ~ Totalh_mg_dose,
      hMG_used == 1 & is.na(Totalh_mg_dose) ~ NA_real_
    )
  )
```

## Trigger
```{r}
print(unique(freshdata$Trigger_med))
```

## Derive baseline total follicles from follicle size counts

```{r}
freshdata <- freshdata %>%
  mutate(
    Baseline_total_follicles = ifelse(
      rowSums(!is.na(select(., starts_with("Follic_cnt")))) == 0,
      NA,
      rowSums(select(., starts_with("Follic_cnt")), na.rm = TRUE)
    )
  )

freshdata <- freshdata %>%
  relocate(
    Baseline_total_follicles,
    .after = Natural_follicle_info
  )
```

```{r}
freshdata %>%
  filter(
    !is.na(No_mature_eggs) &
    !is.na(No_inseminated) &
    No_mature_eggs != No_inseminated
  )

print(sum(freshdata$No_inseminated == 0)) #101
print(sum(freshdata$No_mature_eggs == 0)) #159
```

## first cycle
```{r}
firstcycle <- freshdata %>%
  mutate(
    Cycle_start_date = as.Date(Cycle_start_date)
  ) %>%
  arrange(Unique_number, Cycle_start_date) %>%
  group_by(Unique_number) %>%
  slice_min(order_by = Cycle_start_date, n = 1, with_ties = FALSE) %>%
  ungroup()

print(sum(firstcycle$No_inseminated == 0)) #62 missing
print(sum(firstcycle$No_mature_eggs == 0)) #96 missing 

firstcycle %>%
  filter(
    !is.na(No_mature_eggs) &
    !is.na(No_inseminated) &
    No_mature_eggs != No_inseminated
  )

firstcycle <- firstcycle %>%
  filter(No_inseminated != 0)
```

```{r}
print(sum(is.na(firstcycle$Baseline_AMH))) #266
firstcycle <- firstcycle %>%
  filter(Baseline_AMH != 0)

first
```

```{r}
firstcycle <- firstcycle %>%
  left_join(
    smalldata %>% select(Unique_number, Baseline_total_follicles),
    by = "Unique_number"
  )

firstcycle <- firstcycle %>%
  relocate(
    Baseline_total_follicles.y,
    .after = Natural_follicle_info
  )
```

```{r}
edaparams <- c(
  "Age",
  "BMI",
  "Aetiology",
  "protocol_type_clean",
  "FSH_used",
  "hMG_used",
  "FSH_starting_dose",
  "hMG_starting_dose",
  "Total_fsh_dose",
  "Totalh_mg_dose",
  "Trigger_med",
  "Total_trigger_dose",
  "Final_follicles_lessthan_11.9",
  "Baseline_AMH",
  "Natural_follicle_info",
  "Baseline_total_follicles.y",
  "No_eggs",
  "No_inseminated"
)

firstcycledata <- firstcycle %>%
  all_of(select(edaparams))
```

```{r}
cor.test(firstcycledata$Total_fsh_dose, firstcycledata$No_inseminated, method = "pearson")
```
```{r}
cor.test(firstcycledata$Age, firstcycledata$No_inseminated, method = "pearson")
```
```{r}
cor.test(firstcycledata$Age, firstcycledata$Total_fsh_dose, method = "pearson")
```

```{r}
write_csv(
  firstcycledata,
  here("data", "processed", "freshdata_ext.csv")
)
```

## Just getting first cycle data for the patients to fill in the correct total follicle size: preparing for quick clustering for meeting
```{r}
sum(is.na(firstcycle$Baseline_total_follicles)) #360 missing/471

firstcycle %>%
  count(Unique_number) %>%
  filter(n > 1) #0 all numbers are unique = 471 patients


sum(firstcycle$Unique_number %in% smalldata$Unique_number)

firstcycle_updated <- firstcycle %>%
  left_join(
    smalldata %>% 
      dplyr::select(Unique_number, Baseline_total_follicles),
    by = "Unique_number",
    suffix = c("", "_new")
  ) %>%
  mutate(
    Baseline_total_follicles = dplyr::coalesce(
      Baseline_total_follicles,
      Baseline_total_follicles_new
    )
  ) %>%
  dplyr::select(-Baseline_total_follicles_new)

sum(is.na(firstcycle$Baseline_total_follicles)) #360 missing
sum(is.na(firstcycle$Baseline_AMH)) #314 missing 

sum(is.na(total_data$Baseline_total_follicles)) #607 missing
sum(is.na(firstcycle$Baseline_AMH))
sum(is.na(total_data$Baseline_AMH))

```
