---
title: "CLEAN.1"
output: html_document
date: "2026-02-16"
---

```{r}
library(dplyr)
library(janitor)
library(stringr)
library(purrr)
library(tidyr)
```

```{r}
getwd()
setwd("/Users/kristinastoianova/Desktop/data/")
freshdata <- read.csv("patientdata_fresh1.csv")
frozendata <- read.csv("patientdata_frozen.csv")
```

```{r}
head(freshdata)
#df 521x269
class(freshdata)
```

```{r}
#only 90 have data currently
#removing rows 91-176
freshdata[90, ]
freshdata[91, ]
freshdata[521, ]
freshdata_1 <- freshdata[-c(91:521), ]
```

```{r}
#adds underscores, begins with uppercase letter
freshdata_2 <- freshdata_1 %>%
  clean_names() %>%
  rename_with(~ str_to_title(.))
```

```{r}
#checking how many patients have a second cycle of data entered = 18
sum(!is.na(freshdata_2$Cycle_no_1))
#checking how many patients have a third cycle of data entered = 10
sum(!is.na(freshdata_2$Cycle_no_2))
##does not mean this is their second or third cycle - could be 4th etc.
```

```{r}
#removing data fluid obtained
freshdata_2$Date_fluid_obtained <- NULL
#removing ethnicity, none at the moment
freshdata_2$Ethnicity <- NULL

#removing parameters not of interest - from egg fertilization onwards
#removing from column 57 = ICSI/IVF this is a repeated, unnecessary column
freshdata_2[, 57:ncol(freshdata_2)]

#211/267 parameters will be removed
ncol(freshdata_2) - 56
```

```{r}
#dropping parameters on fertilization/embryo data + future cycles 
freshdata_3 <- freshdata_2[, -c(57:ncol(freshdata_2))]
#90x56 data frame so far
```

## Consented Date
```{r}
sum(str_squish(freshdata_3$Consented_date) == "Not on BBS", na.rm = TRUE) #29 empty fields/missing data
sum(is.na(freshdata_3$Consented_date)) #0 NAs

#replace with NA
freshdata_3$Consented_date[freshdata_3$Consented_date == "Not on BBS"] <- NA
freshdata_3$Consented_date[freshdata_3$Consented_date == "TBC"] <- NA

sum(is.na(freshdata_3$Consented_date)) #30 NA
```

## Create PGD binary column 
```{r}
freshdata_3 <- freshdata_3 %>%
  mutate(
    PGD_binary = if_else(
      str_detect(Aetiology, "PGD"),
      "Y",
      "N"
    ), 
    .after = Aetiology
  )

##Check
freshdata_3 %>%
  filter(
    str_detect(Aetiology, "PGD") &
    PGD_binary != "Y"
  )
#0 violations

colnames(freshdata_3)[colnames(freshdata_3) == "PGD_binary"] <- "PGD"
```

## Aetiology
```{r}
male_terms <- c("Azoospermia", "Azoospermia +", "Male Factor", "Male Factor +", "+ Male Factor", ", Male Factor", "Male Factor Other", "Oligozoospermia", "Oligozoospermia +", "Oligozoospermia and ", "(Oligozoospermia)", "(OAT)", ", Sperm Donor", "(Teratoma)", "(terato)", "()")

for (term in male_terms) {
  freshdata_3$Aetiology <- str_remove_all(
    freshdata_3$Aetiology,
    fixed(term)
  )
} 

#Remove "PGD," or ", PGD"
freshdata_3 <- freshdata_3 %>%
  mutate(
    Aetiology = str_replace_all(Aetiology, "^PGD,\\s*|,\\s*PGD", ""),
    Aetiology = str_replace_all(Aetiology, "\\bPGD\\b", "")
  )

freshdata_3 <- freshdata_3 %>%
  mutate(
    #Replace any remaining commas with +
    Aetiology = str_replace_all(Aetiology, ",\\s*", " + "),
    #Remove duplicate commas and +
    Aetiology = str_replace_all(Aetiology, ",\\s*,", ", "),
    Aetiology = str_replace_all(Aetiology, "\\+\\s*\\+", "+"),
    #Remove leading/trailing commas or +
    Aetiology = str_replace_all(Aetiology, "^[,\\+\\s]+|[,\\+\\s]+$", ""),
    #Trim white space
    Aetiology = str_squish(Aetiology),
    
    #If empty field replace with "No Female Factor"
    Aetiology = if_else(
      Aetiology == "" | is.na(Aetiology),
      "No Female Factor",
      Aetiology
    )
  )

replacements <- c(
  "Fertility Preservation (low ovarian reserve + endometriosis)" = "Diminished Ovarian Reserve + Endometriosis",
  "and DOR" = "Diminished Ovarian Reserve",
  "Researve" = "Reserve",
  "Ovulation Disorder, PCO" = "Ovulation Disorder PCO",
  "Ovulation Disorders/PCO" = "Ovulation Disorder PCO"
)

freshdata_3$Aetiology <- str_replace_all(
  freshdata_3$Aetiology,
  fixed(replacements)
)

freshdata_3$Aetiology[freshdata_3$Aetiology == "Ovulation Disorder + PCO"] <- "Ovulation Disorder PCO"
freshdata_3$Aetiology[newdata$Aetiology == "Same Sex Relatonship"] <- "Same Sex Relationship"
```

## Male Factor
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Male_factor_y_n"] <- "Male_factor"

print(unique(freshdata_3[ ,"Male_factor"])) #N, Y, y
freshdata_3$Male_factor[freshdata_3$Male_factor == "y"] <- "Y"
print(unique(freshdata_3[ ,"Male_factor"]))

#Check that when PGD = Y, "No Female Factor" in Aetiology, Male_factor is "Y"
freshdata_3 %>%
  filter(
    Aetiology == "No Female Factor" &
    Male_factor != "Y" &
    PGD == "Y"
  )
#0 violation
#These are due to PGD
```

## Make PGD and Male Factor column numerical
```{r}
freshdata_3 <- freshdata_3 %>%
  mutate(
    Male_factor = case_when(
      Male_factor == "Y" ~ 1,
      Male_factor == "eN" ~ 0,
      TRUE ~ NA_real_
    )
  )
freshdata_3 <- freshdata_3 %>%
  mutate(
    PGD = case_when(
      PGD == "Y" ~ 1,
      PGD == "N" ~ 0,
      TRUE ~ NA_real_
    )
    )
```

## Funding
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Nhs_pp"] <- "Funding"
print(unique(freshdata_3[ ,"Funding"])) #"PP", "NHS", "NHS ", "PP (self funded)", "missing"

sum(freshdata_3$Funding == "", na.rm = TRUE)
#Replace missing with NA, replace PP (self funded) with PP, Replace "NHS " with NHS

freshdata_3 <- freshdata_3 %>%
  mutate(
    Funding = str_squish(Funding),   #removes trailing spaces
    
    Funding = case_when(
      Funding == "missing" ~ NA,
      Funding == "PP (self funded)" ~ "PP",
      TRUE ~ Funding
    )
  )

print(unique(freshdata_3[ ,"Funding"]))
sum(is.na(freshdata_3$Funding)) #1 missing
```

##Infertility Duration
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Duration_of_infertility_months"] <- "Infertility_duration"

print(unique(freshdata_3[ ,"Infertility_duration"]))

#freshdata$Infertility_duration <- trimws(freshdata$Infertility_duration)

#Non number entries
sum(!grepl("^[0-9]+$", freshdata_3$Infertility_duration))
#45 true numerical entries
#39 missing
sum(freshdata_3$Infertility_duration %in% 
      c("N/A", "n/a", "MISSING", "_", "", "Not on BBS"))
#6 have > signs
sum(grepl("^>", freshdata_3$Infertility_duration), na.rm = TRUE)

#clean
freshdata_3 <- freshdata_3 %>%
  mutate(
    Infertility_duration = str_squish(Infertility_duration),
    Infertility_duration = if_else(
      Infertility_duration %in% c("MISSING", "n/a", "N/A", "", "_", "Not on BBS"),
      NA,
      Infertility_duration
    )
  )

print(unique(freshdata_3[ ,"Infertility_duration"]))
sum(is.na(freshdata_3$Infertility_duration)) #39 missing

unique(freshdata_3$Infertility_duration[grepl("^>", freshdata_3$Infertility_duration)]) ##>24, >48, > 36 --> 2, 3, 4 years

str(freshdata_3$Infertility_duration) #data is chr because of > signs

#remove > signs
freshdata_3 <- freshdata_3 %>%
  mutate(
    Infertility_duration = gsub("^>", "", Infertility_duration)
  )

freshdata_3$Infertility_duration <- as.numeric(freshdata_3$Infertility_duration)
```

#Treatment
```{r}
print(unique(freshdata_3[ ,"Treatment"]))

freshdata_3 <- freshdata_3 %>%
  mutate(
    Treatment = str_squish(Treatment),
    
    Treatment = case_when(
      Treatment %in% c("IVF/ICSI", "ICSI/IVF") ~ "ICSI",
      Treatment == "ICSI-38" ~ "ICSI",
      Treatment == "IVF (35)" ~ "IVF",
      Treatment == "IVF short with pills" ~ "IVF",
      Treatment == "" ~ NA,
      TRUE ~ Treatment
    )
  )

print(unique(freshdata_3[ ,"Treatment"]))
sum(is.na(freshdata_3$Treatment)) #4 NA
```

```{r}
#exploring consented date and date of baseline scan for every individual --> this can be used to extract correct parameters from big data later
dates_qc <- freshdata_3 %>%
  select(Unique_number, Consented_date, Baseline_scan)
```

## DOB and Age
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Dob"] <- "DOB"
colnames(freshdata_3)[colnames(freshdata_3) == "Age_at_present_cycle"] <- "Age"

str(freshdata_3$Age) #num
```

## Baseline hormone levels
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Baseline_fsh"] <- "Baseline_FSH"
sum(freshdata_3$Baseline_FSH == "Not on BBS", na.rm = TRUE) #67 missing values

colnames(freshdata_3)[colnames(freshdata_3) == "Baseline_amh"] <- "Baseline_AMH"
sum(freshdata_3$Baseline_AMH == "Not on BBS", na.rm = TRUE) #43 missing values

colnames(freshdata_3)[colnames(freshdata_3) == "Baseline_afc"] <- "Baseline_AFC"
sum(freshdata_3$Baseline_AFC == "", na.rm = TRUE) #1 missing value
sum(freshdata_3$Baseline_AFC == "Not on BBS", na.rm = TRUE) #63 missing values ##64 total

colnames(freshdata_3)[colnames(freshdata_3) == "Baseline_e2"] <- "Baseline_E2"
sum(freshdata_3$Baseline_E2 == "Not on BBS", na.rm = TRUE) #66 missing values

freshdata_3 <- freshdata_3 %>%
  mutate(
    Baseline_FSH = round(as.numeric(str_remove(Baseline_FSH, "\\s*\\(.*\\)")), 1),
    Baseline_AMH = round(as.numeric(str_remove(Baseline_AMH, "\\s*\\(.*\\)")), 1),
    Baseline_AFC = round(as.numeric(str_remove(Baseline_AFC, "\\s*\\(.*\\)")), 1),
    Baseline_E2  = round(as.numeric(str_remove(Baseline_E2,  "\\s*\\(.*\\)")), 1)
  )

str(freshdata_3$Baseline_FSH) ##numeric
str(freshdata_3$Baseline_AMH) ##numeric
str(freshdata_3$Baseline_AFC) ##numeric
str(freshdata_3$Baseline_E2) ##numeric

sum(is.na(freshdata_3$Baseline_FSH))
```

## Baseline scan
```{r}
print(unique(freshdata_3[ ,"Baseline_scan"]))

freshdata_3 <- freshdata_3 %>%
  mutate(
    Baseline_scan = str_squish(Baseline_scan),
    Baseline_scan = if_else(
      Baseline_scan %in% c("incomplete chart", "_", "N/A"),
      NA,
      Baseline_scan
    )
  )
print(unique(freshdata_3[ ,"Baseline_scan"]))
sum(is.na(freshdata_3$Baseline_scan)) #3 missing
```

##Baseline follicles
```{r}
#Right follicles
colnames(freshdata_3)[colnames(freshdata_3) == "No_of_follicles_on_bl_scam_right"] <- "Baseline_right_follicles"
freshdata_3$Baseline_right_follicles <- trimws(freshdata_3$Baseline_right_follicles)

#6 non number entries, missing data
sum(is.na(freshdata_3$Baseline_right_follicles) |
  !grepl("^\\d+$", freshdata_3$Baseline_right_follicles)
)
#convert to NA
freshdata_3$Baseline_right_follicles[
  !grepl("^\\d+(\\.\\d+)?$", freshdata_3$Baseline_right_follicles)] <- NA
sum(is.na(freshdata_3$Baseline_right_follicles))

str(freshdata_3$Baseline_right_follicles) ##chr
freshdata_3$Baseline_right_follicles <- as.numeric(freshdata_3$Baseline_right_follicles) ##num

#Left follicles 
colnames(freshdata_3)[colnames(freshdata_3) == "No_of_follicles_on_bl_scan_left"] <- "Baseline_left_follicles"

#8 non number entries, missing data
sum(is.na(freshdata_3$Baseline_left_follicles) |
  !grepl("^\\d+$", freshdata_3$Baseline_left_follicles)
)
#convert to NA
freshdata_3$Baseline_left_follicles[
  !grepl("^\\d+(\\.\\d+)?$", freshdata_3$Baseline_left_follicles)] <- NA
sum(is.na(freshdata_3$Baseline_left_follicles))

str(freshdata_3$Baseline_left_follicles) ##chr
freshdata_3$Baseline_left_follicles <- as.numeric(freshdata_3$Baseline_left_follicles) ##num

#Endometrium
colnames(freshdata_3)[colnames(freshdata_3) == "Baseline_endometrium_mm"] <- "Baseline_endometrium"

freshdata_3$Baseline_endometrium <- trimws(freshdata_3$Baseline_endometrium) #removes trailing space
#3 missing values 
sum(freshdata_3$Baseline_endometrium %in% 
  c("_", "N/A", "incomplete chart"),
  na.rm = TRUE
)

#convert to NA
freshdata_3$Baseline_endometrium[
  tolower(freshdata_3$Baseline_endometrium) %in% 
    c("_", "n/a", "incomplete chart")
] <- NA
sum(is.na(freshdata_3$Baseline_endometrium))

str(freshdata_3$Baseline_endometrium) #chr, need to convert to numeric and round to 2dp

#convert to numeric and round to 2dp
freshdata_3$Baseline_endometrium <- round(
  as.numeric(freshdata_3$Baseline_endometrium),
  2
  )
```

## Deriving Total Follicle count
```{r}
#data$Baseline_AFC <- NULL

freshdata_3 <- freshdata_3 %>%
  mutate(Baseline_total_follicles = Baseline_right_follicles + Baseline_left_follicles)

freshdata_3 <- freshdata_3 %>%
  relocate(
    Baseline_total_follicles,
    .after = Baseline_E2
  )
```

## Baseline follicle size
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Size_of_follicle"] <- "Baseline_follicle_size"
unique(freshdata_3$Baseline_follicle_size)

freshdata_3 <- freshdata_3 %>%
  mutate(
    Baseline_follicle_size = str_trim(Baseline_follicle_size),
    #Convert missing data to NA
    Baseline_follicle_size = case_when(
      Baseline_follicle_size %in% c(
        "N/A", "_",
        "see comments ", 
        "see comments",
        "incomplete chart ",
        "incomplete chart",
        "Jun-19"
      ) ~ NA_character_,
      TRUE ~ Baseline_follicle_size
    ),
    #Standardize format
    Baseline_follicle_size = str_replace_all(Baseline_follicle_size, "--", "-"),
    Baseline_follicle_size = str_replace_all(Baseline_follicle_size, "->", ">"),
    Baseline_follicle_size = str_squish(Baseline_follicle_size)
  )
sum(is.na(freshdata_3$Baseline_follicle_size)) #6 missing
table(freshdata_3$Baseline_follicle_size)

##Follicle size function
process_follicle_size <- function(freshdata_3, column, prefix) {
  
  column <- enquo(column)
  
  freshdata_3 %>%
    mutate(
      # Flag >24
      !!paste0(prefix, "_follicle_size_gt24_flag") :=
        as.integer(str_detect(!!column, "^>")),
      
      #Minimum size
      !!paste0(prefix, "_follicle_min_size") :=
        case_when(
          is.na(!!column) ~ NA_real_,
          str_detect(!!column, "^>") ~ 24,
          TRUE ~ as.numeric(str_extract(!!column, "^\\d+"))
        ),
      
      #Maximum size
      !!paste0(prefix, "_follicle_max_size") :=
        case_when(
          is.na(!!column) ~ NA_real_,
          str_detect(!!column, "^>") ~ 25,
          str_detect(!!column, "-") ~ 
            as.numeric(str_extract(!!column, "\\d+$")),
          TRUE ~ as.numeric(str_extract(!!column, "^\\d+"))
        ),
      
      # Spread
      !!paste0(prefix, "_follicle_spread") :=
        !!sym(paste0(prefix, "_follicle_max_size")) -
        !!sym(paste0(prefix, "_follicle_min_size"))
    )
}
```

```{r}
freshdata_3 <- process_follicle_size(
  freshdata_3,
  Baseline_follicle_size,
  "Baseline"
)

freshdata_3 <- freshdata_3 %>%
  relocate(
    Baseline_follicle_min_size,
    Baseline_follicle_max_size,
    Baseline_follicle_size_gt24_flag,
    Baseline_follicle_spread,
    .after = Baseline_follicle_size
  )

freshdata_3$Baseline_follicle_size <- NULL
```

## Protocol
```{r}
print(unique(freshdata_3[ ,"Protocol"])) #19 unique entries
freshdata_3 %>%
  filter(str_detect(Protocol, regex("short", ignore_case = TRUE))) %>%
  pull(Protocol) ##Short antagonist and Short hMG are the same --> rename
```
```{r}
freshdata_3 <- freshdata_3 %>%
  mutate(
    Protocol = str_squish(Protocol),  # remove double spaces
    #Fix 225 iu to 225iu
    Protocol = str_replace_all(Protocol, "(\\d+)\\s+iu", "\\1iu"),
    Protocol = str_replace(Protocol, "hMG\\+rRFSH", "hMG+rFSH"), #typo 
    #Standardize Fertility preservation case ONLY
    Protocol = str_replace(Protocol, "Fertility preservation", "Fertility Preservation")
  )
```

## Split Protocol into 3 columns: Protocol, Drug, Dose
```{r}
freshdata_3 <- freshdata_3 %>%
  mutate(
    #Protocol Type
    Protocol_type = case_when(
      str_detect(Protocol, "Long Agonist") ~ "Long Agonist",
      str_detect(Protocol, "Short Antagonist") ~ "Short Antagonist",
      str_detect(Protocol, "^Short\\b") ~ "Short Antagonist", 
      str_detect(Protocol, "Egg Donor") ~ "Egg Donor",
      str_detect(Protocol, "Fertility Preservation") ~ "Fertility Preservation",
      TRUE ~ NA
    ),
    #Drug administered
    Drug = case_when(
      str_detect(Protocol, "hMG\\+rFSH") ~ "hMG+rFSH",
      str_detect(Protocol, "rFSH") & !str_detect(Protocol, "hMG") ~ "rFSH",
      str_detect(Protocol, "hMG") & !str_detect(Protocol, "rFSH") ~ "hMG",
      TRUE ~ NA
    ),
    
    #Dose 
    Dose_iu = as.numeric(str_extract(Protocol, "\\d+")),
    
    .after = Protocol   #places all new columns after Protocol
  )

#Remove original protocol
freshdata_3$Protocol <- NULL
print(unique(freshdata_3$Protocol_type))

#Remove dose, info comes later
freshdata_3$Dose_iu <- NULL
```

```{r}
#Remove date of additional scan
freshdata_3$Date_of_additional_scan <- NULL
```

## Second scans
```{r}
#Scan day
colnames(freshdata_3)[colnames(freshdata_3) == "Day_of_2nd_scan"] <- "Second_scan_day"

sum(freshdata_3$Second_scan_day == "N/A", na.rm = TRUE) #40 missing values 
freshdata_3$Second_scan_day[tolower(trimws(freshdata_3$Second_scan_day)) == "n/a"] <- NA

#Remove D prefix from day so that it can be treated as numeric
freshdata_3$Second_scan_day <- gsub("^D", "", freshdata_3$Second_scan_day)
str(freshdata_3$Second_scan_day) #chr
freshdata_3$Second_scan_day <- as.numeric(freshdata_3$Second_scan_day) ##num

sum(is.na(freshdata_3$Second_scan_day)) #41
```

```{r}
#Right follicles
colnames(freshdata_3)[colnames(freshdata_3) == "X2nd_scan_no_of_follciles_right"] <- "Second_scan_right_follicles"

sum(freshdata_3$Second_scan_right_follicles == "N/A", na.rm = TRUE) #41 NA
freshdata_3$Second_scan_right_follicles[tolower(trimws(freshdata_3$Second_scan_right_follicles)) == "n/a"] <- NA
sum(is.na(freshdata_3$Second_scan_right_follicles)) #42 missing 

str(freshdata_3$Second_scan_right_follicles) #chr
freshdata_3$Second_scan_right_follicles <- as.numeric(freshdata_3$Second_scan_right_follicles) #num

#Left follicles 
colnames(freshdata_3)[colnames(freshdata_3) == "X2nd_scan_no_of_follciles_left"] <- "Second_scan_left_follicles"

sum(freshdata_3$Second_scan_left_follicles == "N/A", na.rm = TRUE) #41 NA, 1 see comments and 1 _
freshdata_3$Second_scan_left_follicles[tolower(trimws(freshdata_3$Second_scan_left_follicles)) == "n/a"] <- NA
freshdata_3$Second_scan_left_follicles[freshdata_3$Second_scan_left_follicles == "_"] <- NA
freshdata_3$Second_scan_left_follicles[freshdata_3$Second_scan_left_follicles == "see comments"] <- NA
sum(is.na(freshdata_3$Second_scan_left_follicles)) #45 missing

str(freshdata_3$Second_scan_left_follicles) #chr
freshdata_3$Second_scan_left_follicles <- as.numeric(freshdata_3$Second_scan_left_follicles) #num

#Endometrium 
colnames(freshdata_3)[colnames(freshdata_3) == "X2nd_endo_mm"] <- "Second_scan_endometrium"
sum(freshdata_3$Second_scan_endometrium == "N/A", na.rm = TRUE) #40 NA, 1 missing
freshdata_3$Second_scan_endometrium[tolower(trimws(freshdata_3$Second_scan_endometrium)) == "n/a"] <- NA
freshdata_3$Second_scan_endometrium[freshdata_3$Second_scan_endometrium == "missing"] <- NA
sum(is.na(freshdata_3$Second_scan_endometrium)) #42 missing

#chr --> num 2dp
str(freshdata_3$Second_scan_endometrium)
freshdata_3$Second_scan_endometrium <- round(as.numeric(freshdata_3$Second_scan_endometrium), 2)

##Should add units to endometrium (mm) column names 
```

## Follicle Size
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "X2nd_scan_follicle_size"] <- "Second_scan_follicle_size"
unique(freshdata_3$Second_scan_follicle_size)

#clean
freshdata_3 <- freshdata_3 %>%
  mutate(
    Second_scan_follicle_size = str_trim(Second_scan_follicle_size),
    #Convert missing data to NA
    Second_scan_follicle_size = case_when(
      Second_scan_follicle_size %in% c(
        "N/A", "_",
        "see comments ", 
        "see comments",
        "incomplete chart ",
        "incomplete chart",
        "Jun-19"
      ) ~ NA_character_,
      TRUE ~ Second_scan_follicle_size
    ),
    #Standardize format
    Second_scan_follicle_size = str_replace_all(Second_scan_follicle_size, "--", "-"),
    Second_scan_follicle_size = str_replace_all(Second_scan_follicle_size, "->", ">"),
    Second_scan_follicle_size = str_squish(Second_scan_follicle_size)
  )

sum(is.na(freshdata_3$Second_scan_follicle_size)) #41 missing
table(freshdata_3$Second_scan_follicle_size)
```
```{r}
freshdata_3 <- process_follicle_size(
  freshdata_3,
  Second_scan_follicle_size,
  "Second_scan"
)

freshdata_3 <- freshdata_3 %>%
  relocate(
    Second_scan_follicle_min_size,
    Second_scan_follicle_max_size,
    Second_scan_follicle_size_gt24_flag,
    Second_scan_follicle_spread,
    .after = Second_scan_follicle_size
  )

freshdata_3$Second_scan_follicle_size <- NULL
```

## Days of Stimulation
```{r}
print(unique(freshdata_3$Days_of_stim))
str(freshdata_3$Days_of_stim) ##chr
freshdata_3$Days_of_stim <- as.numeric(freshdata_3$Days_of_stim) ##num
sum(is.na(freshdata_3$Days_of_stim)) #1 NA
```

## Date of final scan
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Dx_last_scan_prior_to_trigger"] <- "Date_last_scan_pre_trigger"
print(unique(freshdata_3$Date_last_scan_pre_trigger))
freshdata_3$Date_last_scan_pre_trigger <- NULL

#str(freshdata_3$Date_last_scan_pre_trigger)
#freshdata_3$Date_last_scan_pre_trigger[freshdata_3$Date_last_scan_pre_trigger == "incomplete chart "] <- NA
#freshdata_3$Date_last_scan_pre_trigger[freshdata_3$Date_last_scan_pre_trigger == ""] <- NA
```

## Other (drug)
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Other"] <- "Other_drug"
print(unique(freshdata_3[ ,"Other_drug"]))
```
## Last day of stim
```{r}
freshdata_3$Last_day_of_stim <- gsub("^D", "", freshdata_3$Last_day_of_stim)
str(freshdata_3$Last_day_of_stim) ##chr --> num
freshdata_3$Last_day_of_stim <- as.numeric(freshdata_3$Last_day_of_stim)
```

## Day of trigger
```{r}
freshdata_3$Day_of_trigger <- gsub("^D", "", freshdata_3$Day_of_trigger)
freshdata_3$Day_of_trigger <- as.numeric(freshdata_3$Day_of_trigger)
str(freshdata_3$Day_of_trigger) ##num
sum(is.na(freshdata_3$Day_of_trigger)) #8 missing
```

```{r}
trigger_qc <- freshdata_3 %>%
  select(Unique_number, Last_day_of_stim, Day_of_trigger)
```

## Day of Egg Collection (EC)
```{r}
freshdata_3$Day_of_ec <- gsub("^D", "", freshdata_3$Day_of_ec)
sum(is.na(freshdata_3$Day_of_EC)) #0
```

## Date and Day of EC
```{r}
#freshdata_3$Date_of_ec[freshdata_3$Date_of_ec == ""] <- NA
#sum(is.na(freshdata_3$Date_of_ec)) ##37 NA
freshdata_3$Date_of_ec <- NULL

freshdata_3$Day_of_ec <- as.numeric(freshdata_3$Day_of_ec)
str(freshdata_3$Day_of_ec) #num
```

```{r}
str(freshdata_3$No_eggs) #int
str(freshdata_3$No_mature_eggs) #int
```

## Final Scans
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Follicles_right"] <- "Final_follicles_right"
print(unique(freshdata_3$Final_follicles_right))
freshdata_3$Final_follicles_right <- as.numeric(freshdata_3$Final_follicles_right)
sum(is.na(freshdata_3$Final_follicles_right)) #1 NA

colnames(freshdata_3)[colnames(freshdata_3) == "Follicles_left"] <- "Final_follicles_left"
print(unique(freshdata_3$Final_follicles_left))
freshdata_3$Final_follicles_left <- as.numeric(freshdata_3$Final_follicles_left)
sum(is.na(freshdata_3$Final_follicles_left)) #4 NA

colnames(freshdata_3)[colnames(freshdata_3) == "Endo_at_ec_mm"] <- "Final_endometrium"
freshdata_3$Final_endometrium <- round(as.numeric(freshdata_3$Final_endometrium), 2)
sum(is.na(freshdata_3$Final_endometrium)) #2 NA
```

## Final Follicle size
```{r}
colnames(freshdata_3)[colnames(freshdata_3) == "Foll_size_mm"] <- "Final_follicle_size"

unique(freshdata_3$Final_follicle_size)

#clean
freshdata_3 <- freshdata_3 %>%
  mutate(
    Final_follicle_size = str_trim(Final_follicle_size),
    #Convert missing data to NA
    Final_follicle_size = case_when(
      Final_follicle_size %in% c(
        "N/A", "_",
        "see comments ", 
        "see comments",
        "incomplete chart ",
        "incomplete chart",
        "Jun-19"
      ) ~ NA_character_,
      TRUE ~ Final_follicle_size
    ),
    #Standardize format
    Final_follicle_size = str_replace_all(Final_follicle_size, "--", "-"),
    Final_follicle_size = str_replace_all(Final_follicle_size, "->", ">"),
    Final_follicle_size = str_squish(Final_follicle_size)
  )

sum(is.na(freshdata_3$Final_follicle_size)) #41 missing
table(freshdata_3$Final_follicle_size)
```
```{r}
freshdata_3 <- process_follicle_size(
  freshdata_3,
  Final_follicle_size,
  "Final"
)

freshdata_3 <- freshdata_3 %>%
  relocate(
    Final_follicle_min_size,
    Final_follicle_max_size,
    Final_follicle_size_gt24_flag,
    Final_follicle_spread,
    .after = Final_follicle_size
  )

freshdata_3$Final_follicle_size <- NULL
```

## Removing sperm data
```{r}
freshdata_3$Sperm_source_f_fz <- NULL

freshdata_3$Pre_wash_sperm_concentration_mil_ml <- NULL
freshdata_3$Pre_wash_sperm_motility <- NULL
freshdata_3$Pre_wash_sperm_morphology_normal_forms <- NULL

freshdata_3$Post_wash_sperm_concentration_mil_ml <- NULL
freshdata_3$Post_wash_sperm_motility <- NULL
freshdata_3$Semen_volume_ml <- NULL

freshdata_3$Sperm_prep_dg_mini_none <- NULL
freshdata_3$Sperm_mobil_y_n <- NULL
```

## Cycle number --> removing patients whose first entered data is not their first cycle and frozen data
```{r}
print(unique(freshdata_3$Cycle_no))
sum(freshdata_3$Cycle_no == "1") #78
sum(freshdata_3$Cycle_no == "2") #8
sum(freshdata_3$Cycle_no == "3") #3
sum(freshdata_3$Cycle_no == "4") #1

str(freshdata_3$Cycle_no) #int

freshdata_3 <- freshdata_3 %>%
  filter(Cycle_no == 1)

print(unique(freshdata_3$Treatment))
sum(freshdata_3$Cycle_no == "1")

```

```{r}
freshdata_3$H_mg <- NULL
freshdata_3$DOB <- NULL
freshdata_3$Consented_date <- NULL
freshdata_3$Baseline_scan <- NULL
freshdata_3$Drug <- NULL
```
```{r}
#freshdata_3 <- freshdata_3 %>% select(-H_mg)
```

```{r}
write.csv(freshdata_3, "cleaned_freshdata.csv", row.names = FALSE)
```

