---
title: "Clustering"
output: html_document
date: "2026-03-22"
---

```{r setup, message=FALSE, warning=FALSE}
library(naniar)
library(tidyverse)
library(gridExtra)
library(corrplot)
library(pheatmap)
library(cluster)
library(factoextra)
library(dplyr)
library(here)
here::here()

freshdata <- read_csv(
  here("data","processed","freshdata_cleaned1.csv"),
  name_repair = "minimal"
)
```

## Encode and Group Aetiology
```{r}
print(unique(freshdata$Aetiology))

freshdata <- freshdata %>%
  mutate(Aetiology = case_when(
    
    #Female factor group
    Aetiology %in% c("Tubal Factors", 
                     "DOR",
                     "Endometriosis",
                     "Uterine Factor",
                     "Endometriosis + Uterine Factor",
                     "Endometriosis + Tubal Factors",
                     "PCO",
                     "DOR + Endometriosis + PCO")  ~ "Female_factor",
    
    #Other group
    Aetiology %in% c("Other",
                     "Same Sex Relationship",
                     "Fertility Preservation",
                     "Single woman",        # lowercase w from your data
                     "Surrogate")           ~ "Other",
    
    #Unexplained
    Aetiology == "Unexplained"              ~ "Unexplained",
    
    #No female factor
    Aetiology %in% c("No Female Factor",   # capital F from your data
                     "No female factor")   ~ "No_female_factor",
  ))


# Numeric encoding
freshdata <- freshdata %>%
  mutate(Aetiology = case_when(
    Aetiology == "Female_factor"    ~ 1,
    Aetiology == "Unexplained"      ~ 2,
    Aetiology == "Other"            ~ 3,
    Aetiology == "No_female_factor" ~ 4,
    TRUE ~ NA_real_
  ))

str(freshdata$Aetiology)


##convert back to labels as not good to have 1 to 4, should be one hot encoded if used!
freshdata$Aetiology <- factor(
  freshdata$Aetiology,
  levels = c(1,2,3,4),
  labels = c("Female_factor", "Unexplained", "Other", "No_female_factor")
)
```

## Prepare --> extract first cycles only (78)
```{r}
filter_first_cycle <- function(df){
  df %>% filter(Cycle_no == 1)
}

freshdata <- freshdata %>%
  filter_first_cycle()
```

## Extract clustering variables and prepare: 
```{r}
cluster_vars <- c("Aetiology", "Age", "BMI", "Baseline_AMH", "Baseline_total_follicles", "GnRH_antagonist_duration", "GnRH_antagonist_total_dose", "FSH_total_dose", "FSH_duration", "hMG_duration", "hMG_total_dose", "Last_day_of_stim")

cluster_data <- freshdata[, cluster_vars] %>%
  mutate(across(everything(),
                ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

##SCALE - produces matrix
cluster_scaled <- scale(cluster_data)
```

## Choose optimal k 
```{r}
wss <- sapply(1:8, function(k) {
  pam(cluster_scaled, k)$objective[1]
})

##Elbow plot suggests 3 clusters is optimal
plot(1:8, wss, type = "b", pch = 19, col = "#1D9E75",
     xlab = "Number of clusters (k)",
     ylab = "Total within-cluster distance",
     main = "Elbow method — optimal k")
```
```{r}
avg_sil <- sapply(2:8, function(k) {
  p <- pam(cluster_scaled, k)
  s <- silhouette(p$clustering, dist(cluster_scaled))
  mean(s[, 3])
})

plot(2:8, avg_sil, type = "b", pch = 19, col = "#7F77DD",
     xlab = "Number of clusters (k)",
     ylab = "Average silhouette width",
     main = "Silhouette method — optimal k")

k <- which.max(avg_sil) + 1
cat("Optimal k:", k, "\n")
```
```{r}
#Set k and run PAM
k <- 3
set.seed(42)
pam_result <- pam(cluster_scaled, k)

cat("\nCluster sizes:\n")
print(table(pam_result$clustering))
```

```{r}
sil <- silhouette(pam_result$clustering, dist(cluster_scaled))
plot(sil,
     col = c("#1D9E75","#7F77DD","#D85A30"),
     border = NA,
     main = "Silhouette plot — k=3")
cat("Average silhouette width:", round(mean(sil[, 3]), 3), "\n")
```

## Add clusters back to dataset
```{r}
freshdata$cluster <- as.factor(pam_result$clustering)
```

## Cluster interpretation --> use outcome: number of mature eggs
```{r}
cluster_summary <- freshdata %>%
  group_by(cluster) %>%
  summarise(
    n                  = n(),
    median_mature_eggs = median(No_mature_eggs, na.rm = TRUE),
    IQR_mature_eggs    = IQR(No_mature_eggs, na.rm = TRUE),
    median_age         = median(Age, na.rm = TRUE),
    median_AMH         = median(Baseline_AMH, na.rm = TRUE),
    median_total_follicles         = median(Baseline_total_follicles, na.rm = TRUE),
    median_FSH_dose    = median(FSH_total_dose, na.rm = TRUE),
    median_hMG_dose    = median(hMG_total_dose, na.rm = TRUE),
    median_GnRH_dose   = median(GnRH_antagonist_total_dose, na.rm = TRUE)
  )
print(cluster_summary)
```

```{r}
clusplot(cluster_scaled,
         pam_result$clustering,
         color = TRUE,
         col.p = c("#1D9E75","#7F77DD","#D85A30")[pam_result$clustering],
         col.clus = c("#1D9E75","#7F77DD","#D85A30"),
         shade = FALSE,
         lines = 0,
         main = "PAM clustering — k=3",
         sub = "")
```

```{r}
library(ggplot2)
ggplot(freshdata,
       aes(x = cluster, y = No_mature_eggs, fill = cluster)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.4) +
  geom_jitter(width = 0.15, alpha = 0.3) +
  scale_fill_manual(values = c("#9FE1CB","#CECBF6","#F5C4B3")) +
  labs(title = "Mature oocytes by cluster",
       x = "Cluster", y = "No. mature oocytes") +
  theme_minimal()
```
```{r}
kruskal.test(No_mature_eggs ~ cluster, data = freshdata)
```
```{r}
summary(freshdata$Baseline_AMH)
summary(freshdata$Baseline_total_follicles)
```
## Only patient data

```{r}
cluster_vars1 <- c("Age", "BMI", "Baseline_AMH", "Baseline_total_follicles")

cluster_data1 <- freshdata[, cluster_vars1] %>%
  mutate(across(everything(),
                ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

##SCALE - produces matrix
cluster_scaled1 <- scale(cluster_data1)
```

```{r}
## Choose optimal k 
wss1 <- sapply(1:8, function(k) {
  pam(cluster_scaled1, k)$objective[1]
})

##Elbow plot suggests 3 clusters is optimal
plot(1:8, wss1, type = "b", pch = 19, col = "#1D9E75",
     xlab = "Number of clusters (k)",
     ylab = "Total within-cluster distance",
     main = "Elbow method — optimal k")
```

```{r}
avg_sil1 <- sapply(2:8, function(k) {
  p <- pam(cluster_scaled1, k)
  s <- silhouette(p$clustering, dist(cluster_scaled1))
  mean(s[, 3])
})

plot(2:8, avg_sil1, type = "b", pch = 19, col = "#7F77DD",
     xlab = "Number of clusters (k)",
     ylab = "Average silhouette width",
     main = "Silhouette method — optimal k")

k <- which.max(avg_sil1) + 1
cat("Optimal k:", k, "\n")
```

```{r}
#Set k and run PAM
k <- 2
set.seed(42)
pam_result1 <- pam(cluster_scaled1, k)

cat("\nCluster sizes:\n")
print(table(pam_result1$clustering))
```
```{r}
sil1 <- silhouette(pam_result1$clustering, dist(cluster_scaled1))
plot(sil1,
     col = c("#1D9E75","#7F77DD","#D85A30"),
     border = NA,
     main = "Silhouette plot — k=2")
cat("Average silhouette width:", round(mean(sil[, 2]), 2), "\n")
```

```{r}
freshdata$cluster1 <- as.factor(pam_result1$clustering)

cluster_summary1 <- freshdata %>%
  group_by(cluster1) %>%
  summarise(
    n                  = n(),
    median_mature_eggs = median(No_mature_eggs, na.rm = TRUE),
    median_age         = median(Age, na.rm = TRUE),
    median_AMH         = median(Baseline_AMH, na.rm = TRUE),
    median_total_follicles         = median(Baseline_total_follicles, na.rm = TRUE),
    median_BMI         = median(BMI, na.rm = TRUE)
  )
print(cluster_summary1)
```
```{r}
clusplot(cluster_scaled1,
         pam_result1$clustering,
         color = TRUE,
         col.p = c("#1D9E75","#7F77DD","#D85A30")[pam_result1$clustering],
         col.clus = c("#1D9E75","#7F77DD","#D85A30"),
         shade = FALSE,
         lines = 0,
         main = "PAM clustering — k=2",
         sub = "")
```
```{r}
ggplot(freshdata,
       aes(x = cluster1, y = No_mature_eggs, fill = cluster1)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.4) +
  geom_jitter(width = 0.15, alpha = 0.3) +
  scale_fill_manual(values = c("#9FE1CB","#CECBF6","#F5C4B3")) +
  labs(title = "Mature oocytes by cluster",
       x = "Cluster", y = "No. mature oocytes") +
  theme_minimal()
```
```{r}
wilcox.test(No_mature_eggs ~ cluster1, data = freshdata)
```
## Exploring aetiology in clusters
```{r}
table(freshdata$cluster1, freshdata$Aetiology)
prop.table(table(freshdata$cluster1, freshdata$Aetiology), 1)
```
```{r}
freshdata %>%
  ggplot(aes(x = cluster1, fill = Aetiology)) +
  geom_bar(position = "fill") +
  labs(
    y = "Proportion",
    title = "Distribution of aetiology across clusters"
  ) +
  theme_minimal()
```

```{r}
chisq.test(table(freshdata$cluster1, freshdata$Aetiology)) #Aetiology is not associated with a cluster
```
```{r}
treatment_by_cluster1 <- freshdata %>%
  filter(!is.na(cluster1)) %>%
  group_by(cluster1) %>%
  summarise(
    n = n(),
    median_FSH_dose = median(FSH_total_dose, na.rm = TRUE),
    IQR_FSH_dose = IQR(FSH_total_dose, na.rm = TRUE),
    median_FSH_duration = median(FSH_duration, na.rm = TRUE),
    median_hMG_dose = median(hMG_total_dose, na.rm = TRUE),
    median_mature_eggs = median(No_mature_eggs, na.rm = TRUE),
    median_AMH = median(Baseline_AMH, na.rm = TRUE),
    median_baseline_total_follicles = median(Baseline_total_follicles, na.rm = TRUE),
    median_GnRH_antagonist_dose = median(GnRH_antagonist_total_dose, na.rm = TRUE),
    median_GnRH_antagonist_duration = median(GnRH_antagonist_duration, na.rm = TRUE),
    median_age = median(Age, na.rm = TRUE)
  )

treatment_by_cluster1
```

```{r}
wilcox.test(FSH_total_dose ~ cluster1, data = freshdata)
wilcox.test(FSH_duration ~ cluster1, data = freshdata)
wilcox.test(hMG_total_dose ~ cluster1, data = freshdata)
```
```{r}
ggplot(freshdata %>% filter(!is.na(cluster1)),
       aes(x = cluster1, y = FSH_total_dose, fill = cluster1)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.4) +
  geom_jitter(width = 0.15, alpha = 0.3) +
  labs(title = "FSH total dose by biological cluster",
       x = "Cluster", y = "FSH total dose") +
  theme_minimal()
```
```{r}
ggplot(freshdata %>% filter(!is.na(cluster1)),
       aes(x = cluster1, y = hMG_total_dose, fill = cluster1)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.4) +
  geom_jitter(width = 0.15, alpha = 0.3) +
  labs(title = "hMG total dose by biological cluster",
       x = "Cluster", y = "hMG total dose") +
  theme_minimal()
```

##Does high FSH in high-AMH patients (cluster 2) linked to over-responsiveness

```{r}
freshdata <- freshdata %>%
  mutate(
    over_responsive = ifelse(No_mature_eggs >= 20, 1, 0),
    low_response = ifelse(No_mature_eggs < 5, 1, 0)
  )
```

```{r}
ggplot(
  freshdata %>% filter(cluster1 == 2),
  aes(x = factor(over_responsive), y = FSH_total_dose, fill = factor(over_responsive))
) +
  geom_boxplot(alpha = 0.8) +
  labs(title = "FSH dose and over-response in high-reserve cluster",
       x = "Over-response (0 = no, 1 = yes)",
       y = "FSH total dose") +
  theme_minimal()
```
```{r}
high_cluster_data <- freshdata %>%
  filter(cluster1 == 2) %>%
  select(over_responsive, FSH_total_dose, Age, Baseline_AMH, Baseline_total_follicles, BMI) %>%
  na.omit()

model_over <- glm(
  over_responsive ~ FSH_total_dose + Age + Baseline_AMH + Baseline_total_follicles + BMI,
  data = high_cluster_data,
  family = binomial()
)

summary(model_over)
```

## Does FSH improve outcomes in lower AMH patients
```{r}
ggplot(freshdata %>% filter(cluster1 == 1),
       aes(x = FSH_total_dose, y = No_mature_eggs)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = TRUE) +
  labs(title = "FSH dose vs mature oocytes in lower-reserve cluster",
       x = "FSH total dose", y = "No. mature oocytes") +
  theme_minimal()
```
```{r}
low_cluster_data <- freshdata %>%
  filter(cluster1 == 1) %>%
  select(No_mature_eggs, FSH_total_dose, Age, Baseline_AMH, Baseline_total_follicles, BMI) %>%
  na.omit()

model_low <- lm(
  No_mature_eggs ~ FSH_total_dose + Age + Baseline_AMH + Baseline_total_follicles + BMI,
  data = low_cluster_data
)

summary(model_low)
```
## Interaction effect in whole dataset - does the effect of FSH differ depending on reserve? Does the relationship between FSH and mature oocytes change depending on AMH
```{r}
library(MASS)

interaction_data <- freshdata %>%
  select(No_mature_eggs, FSH_total_dose, Baseline_AMH, Baseline_total_follicles, Age, BMI) %>%
  na.omit()

model_interaction <- glm.nb(
  No_mature_eggs ~ FSH_total_dose * Baseline_AMH +
    Baseline_total_follicles + Age + BMI,
  data = interaction_data
)

summary(model_interaction)
```

## Interaction effect per cluster

```{r}
interaction_cluster_data <- freshdata %>%
  filter(FSH_total_dose > 0, !is.na(cluster1)) %>%
  dplyr::select(
    No_mature_eggs,
    FSH_total_dose,
    cluster1,
    Baseline_total_follicles,
    Baseline_AMH,
    Age,
    BMI
  ) %>%
  na.omit()

model_cluster_interaction <- glm.nb(
  No_mature_eggs ~ FSH_total_dose * cluster1 +
    Baseline_total_follicles + Baseline_AMH + Age + BMI,
  data = interaction_cluster_data
)

summary(model_cluster_interaction)
```

