---
title: "Distributions and Descriptive Stats"
output: html_document
date: "2026-04-27"
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggcorrplot)
library(patchwork)
library(corrplot)

library(here)
here::here()

shortdata <- read_csv(
  here("data", "processed", "shortprotocol_firstconsentedcycle.csv"),
  name_repair = "minimal"
)
```

# Structural zeros:
# If a medication was not used, the dose/duration value is not a true biological 0
# It should be treated as NA in the exposed-only variable

```{r}
make_real_vars <- function(data, used_col, value_cols) {
  
  used_col <- enquo(used_col)
  
  data %>%
    mutate(
      across(
        all_of(value_cols),
        ~ if_else(!!used_col == 1, .x, NA_real_),
        .names = "{.col}_real"
      )
    )
}

FSH_vars <- c(
  "FSH_initial_dose",
  "FSH_total_dose",
  "FSH_final_dose",
  "FSH_min_dose",
  "FSH_max_dose",
  "FSH_duration",
  "FSH_dose_changes",
  "FSH_avg_daily"
)

hMG_vars <- c(
  "hMG_initial_dose",
  "hMG_total_dose",
  "hMG_final_dose",
  "hMG_min_dose",
  "hMG_max_dose",
  "hMG_duration",
  "hMG_dose_changes",
  "hMG_avg_daily"
)

shortdata <- shortdata %>%
  make_real_vars(FSH_used, FSH_vars) %>%
  make_real_vars(hMG_used, hMG_vars) %>%
  make_real_vars(hCG_trigger_used, "hCG_trigger_dose") %>%
  make_real_vars(GnRH_agonist_trigger_used, "GnRH_agonist_trigger_dose")

shortdata <- shortdata %>%
  mutate(
    FSH_only = if_else(Treatment_type == "FSH_only", 1, 0),
    hMG_only = if_else(Treatment_type == "hMG_only", 1, 0),
    Dual     = if_else(Treatment_type == "Dual", 1, 0)
  )
```

## Manual change to dual trigger patient

```{r}
# 2313544U
shortdata[shortdata$Unique_number == "2313544U", c("hCG_trigger_used", "GnRH_agonist_trigger_used", "hCG_trigger_dose_real","GnRH_agonist_trigger_dose_real")] <- 
  list(1, 1, 250, 50)
```

## Pre function dealing with structural 0s

```{r}
##initial dose
shortdata <- shortdata %>%
  mutate(
    FSH_initial_dose_real = if_else(FSH_used == 1, FSH_initial_dose, NA_real_),
    hMG_initial_dose_real = if_else(hMG_used == 1, hMG_initial_dose, NA_real_)
  )

## total dose
shortdata <- shortdata %>%
  mutate(
    FSH_total_dose_real = if_else(FSH_used == 1, FSH_total_dose, NA_real_),
    hMG_total_dose_real = if_else(hMG_used == 1, hMG_total_dose, NA_real_)
  )

shortdata <- shortdata %>%
  mutate(
    # Final dose
    FSH_final_dose_real = if_else(FSH_used == 1, FSH_final_dose, NA_real_),
    hMG_final_dose_real = if_else(hMG_used == 1, hMG_final_dose, NA_real_),
    # Minimum dose
    FSH_min_dose_real = if_else(FSH_used == 1, FSH_min_dose, NA_real_),
    hMG_min_dose_real = if_else(hMG_used == 1, hMG_min_dose, NA_real_),
    # Maximum dose
    FSH_max_dose_real = if_else(FSH_used == 1, FSH_max_dose, NA_real_),
    hMG_max_dose_real = if_else(hMG_used == 1, hMG_max_dose, NA_real_),
    # Duration
    FSH_duration_real = if_else(FSH_used == 1, FSH_duration, NA_real_),
    hMG_duration_real = if_else(hMG_used == 1, hMG_duration, NA_real_),
    # Dose changes
    FSH_dose_changes_real = if_else(FSH_used == 1, FSH_dose_changes, NA_real_),
    hMG_dose_changes_real = if_else(hMG_used == 1, hMG_dose_changes, NA_real_),
    # Average daily dose
    FSH_avg_daily_real = if_else(FSH_used == 1, FSH_avg_daily, NA_real_),
    hMG_avg_daily_real = if_else(hMG_used == 1, hMG_avg_daily, NA_real_)
  )

shortdata <- shortdata %>%
  mutate(
    hCG_trigger_dose_real = if_else(hCG_trigger_used == 1, hCG_trigger_dose, NA_real_),
    GnRH_trigger_dose_real = if_else(GnRH_agonist_trigger_used == 1, GnRH_agonist_trigger_dose, NA_real_)
  )
```

## Descriptive Stats

```{r}
descriptive_stats <- function(data, vars) {
  
  data %>%
    select(all_of(vars)) %>%
    pivot_longer(
      cols = everything(),
      names_to = "Variable",
      values_to = "Value"
    ) %>%
    group_by(Variable) %>%
    summarise(
      n_available = sum(!is.na(Value)),
      missing_n = sum(is.na(Value)),
      missing_percent = round(100 * missing_n / n(), 1),
      mean = round(mean(Value, na.rm = TRUE), 2),
      SD = round(sd(Value, na.rm = TRUE), 2),
      median = round(median(Value, na.rm = TRUE), 2),
      Q1 = round(quantile(Value, 0.25, na.rm = TRUE), 2),
      Q3 = round(quantile(Value, 0.75, na.rm = TRUE), 2),
      IQR = round(IQR(Value, na.rm = TRUE), 2),
      min = round(min(Value, na.rm = TRUE), 2),
      max = round(max(Value, na.rm = TRUE), 2),
      .groups = "drop"
    )
}

vars_to_summarise <- c(
  "Age", "BMI", "Baseline_AMH", "Baseline_total_follicles", "Baseline_endometrium", "FSH_initial_dose_exposed", "hMG_initial_dose_exposed", "Total_gonadotropin_dose", "Starting_gonadotropin_dose", "GnRH_antagonist_total_dose", "GnRH_antagonist_duration", "Final_total_follicles", "Final_follicles_lessthan_11.9", "Final_endometrium_ec", "GnRH_trigger_dose_exposed", "hCG_trigger_dose_exposed", "No_eggs", "No_mature_eggs", "No_fertilised"
)

summary_table <- descriptive_stats(shortdata, vars_to_summarise)
summary_table
```

### CORRELATION MATRICES

## Names to go on correlation matrices for tidy look
```{r}
pretty_label_map <- c(
  "Age" = "Age",
  "BMI" = "BMI",
  "Baseline_AMH" = "Baseline AMH",
  "Baseline_total_follicles" = "Baseline follicles",
  "Baseline_endometrium" = "Baseline endometrium",
  
  # Overall stimulation variables
  "FSH_used" = "FSH used",
  "hMG_used" = "hMG used",
  "Starting_gonadotropin_dose" = "Starting gonadotropin dose",
  "Total_gonadotropin_dose" = "Total gonadotropin dose",
  
  # FSH variables
  "FSH_initial_dose_real" = "FSH starting dose",
  "FSH_total_dose_real" = "Total FSH dose",
  "FSH_final_dose_real" = "FSH final dose",
  "FSH_min_dose_real" = "FSH min dose",
  "FSH_max_dose_real" = "FSH max dose",
  "FSH_duration_real" = "FSH duration",
  "FSH_dose_changes_real" = "FSH dose changes",
  "FSH_avg_daily_real" = "Mean daily FSH dose",
  
  # hMG variables
  "hMG_initial_dose_real" = "hMG starting dose",
  "hMG_total_dose_real" = "Total hMG dose",
  "hMG_final_dose_real" = "hMG final dose",
  "hMG_min_dose_real" = "hMG minimum dose",
  "hMG_max_dose_real" = "hMG maximum dose",
  "hMG_duration_real" = "hMG duration",
  "hMG_dose_changes_real" = "hMG dose changes",
  "hMG_avg_daily_real" = "Mean daily hMG dose",
  
  "FSH_only" = "FSH only",
  "hMG_only" = "hMG only",
  "Dual" = "Dual",
  
  ## GnRH antagonist
  "GnRH_antagonist_total_dose" = "GnRH antagonist total dose",
  "GnRH_antagonist_duration" = "GnRH antagonist duration",
  
  "Last_day_of_stim" = "Duration of stim",
  
  ## Trigger
  "hCG_trigger_dose_real" = "hCG trigger dose",
  "GnRH_agonist_trigger_dose_real" = "GnRH agonist trigger dose",
  
  ## Male
  
  "Pre_wash_sperm_conc" = "Pre wash sperm conc.",
  "Pre_wash_sperm_motility" = "Pre wash sperm motility",
  "Pre_wash_normal_sperm_morphology" = "Pre wash normal sperm morph",
  "Post_wash_sperm_conc" = "Post wash sperm conc",
  "Post_wash_sperm_motility" = "Post wash sperm motility",
  "Semen_vol" = "Semen vol",
  
  # Outcomes
  "No_eggs" = "Total oocytes",
  "No_mature_eggs" = "MII oocytes",
  "No_fertilised" = "Fertilised oocytes",
  "Fertilisation_rate" = "Fertilisation rate",
  "Proportion_mature" = "Proportion mature",
  
  #
  "Final_follicles_lessthan_11.9" = "Final follicles\n<11.9mm",
  "Final_total_follicles" = "Final follicles",
  "Final_endometrium_ec" = "Final endometrium"
)

apply_pretty_names <- function(corr_matrix, label_map = pretty_label_map) {
  original_names <- colnames(corr_matrix)
  pretty_names <- ifelse(
    original_names %in% names(label_map),
    label_map[original_names],
    gsub("_", " ", original_names)
  )
  colnames(corr_matrix) <- pretty_names
  rownames(corr_matrix) <- pretty_names
  return(corr_matrix)
}
```

## CORRELATIONS

## 1. Baseline ~ Outcome
```{r, fig.width=6, fig.height=7}
corr_vars1 <- c("Age", "BMI", "Baseline_AMH", "Baseline_total_follicles", "Baseline_endometrium", "No_mature_eggs")

corr_data1 <- shortdata %>%
  dplyr::select(all_of(corr_vars1))

corr_matrix1 <- cor(
  corr_data1,
  use = "pairwise.complete.obs",
  method = "spearman"
)

corr_matrix1_pretty <- apply_pretty_names(corr_matrix1)

corrplot1 <- corrplot(
  corr_matrix1_pretty,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  addCoef.col = "black",
  number.cex = 0.7,
  tl.cex = 0.8,
  tl.col = "black",
  diag = FALSE
)

corrplot1 <- ggcorrplot(
  corr_matrix1_pretty,
  type = "lower",
  lab = TRUE,
  lab_size = 5,
  hc.order = TRUE,
  hc.method = "ward.D2",
  colors = c("#2166AC", "white", "#D6604D"),
  ggtheme = theme_void()
) +
  scale_fill_gradientn(
    colours = c("#2166AC", "white", "#D6604D"),
    limits = c(-1, 1),
    name = "Spearman r"
  ) +
  #labs(
   # title = "Baseline and Oocyte Variables"
  #) +
  theme(
    plot.title = element_text(
      size = 20,
      face = "bold",
      hjust = 0.5
    ),
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1,
      size = 16,
      #face = "bold",
      colour = "black"
    ),
    axis.text.y = element_text(
      size = 20,
      #face = "bold",
      colour = "black"
    ),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12),
    plot.margin = margin(5, 5, 5, 5)
  )

corrplot1

ggsave(
  "corrplot1.png",
  plot = corrplot1,
  #width = 16,
 # height = 9.5,
 # units = "cm",
  bg = "white"
)


ggsave("corrplot1.png", corrplot1, dpi = 600)
```

## 2. Baseline ~ Protocol Decisions

```{r, fig.width=12, fig.height=8}
corr_vars2 <- c("Starting_gonadotropin_dose", "Total_gonadotropin_dose",
  "FSH_initial_dose_real",
  "FSH_total_dose_real",
  "FSH_final_dose_real",
  "FSH_min_dose_real",
  "FSH_max_dose_real",
  "FSH_duration_real",
  "FSH_avg_daily_real",
  "hMG_initial_dose_real",
  "hMG_total_dose_real",
  "hMG_final_dose_real",
  "hMG_min_dose_real",
  "hMG_max_dose_real",
  "hMG_duration_real",
  "hMG_avg_daily_real",
  "Last_day_of_stim",
  "Final_total_follicles",
  "Final_follicles_lessthan_11.9",
  "No_eggs",
  "No_mature_eggs"
)

##little variation in trigger doses so exclude
##exclude FSH dose change as not many

corr_data2 <- shortdata %>%
  dplyr::select(all_of(corr_vars2))

## remove columns with too few variations
corr_data2_clean <- corr_data2 %>%
  dplyr::select(where(~ sum(!is.na(.)) >= 3)) %>%
  dplyr::select(where(~ n_distinct(., na.rm = TRUE) > 1))

corr_matrix2 <- cor(
  corr_data2_clean,
  use = "pairwise.complete.obs",
  method = "spearman"
)

corr_matrix2 <- cor(
  corr_data2,
  use = "pairwise.complete.obs",
  method = "spearman"
)

corr_matrix2_pretty <- apply_pretty_names(corr_matrix2)

corrplot(
  corr_matrix2_pretty,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  #addCoef.col = "black",
  number.cex = 0.5,
  tl.cex = 0.8,
  tl.col = "black",
  diag = FALSE
)

corrplot2 <- ggcorrplot(
  corr_matrix2_pretty,
  type = "lower",
  #lab = TRUE,
  lab_size = 3.5,
  hc.order = TRUE,
  hc.method = "ward.D2",
  colors = c("#2166AC", "white", "#D6604D"),
  ggtheme = theme_void()
) +
  scale_fill_gradientn(
    colours = c("#2166AC", "white", "#D6604D"),
    limits = c(-1, 1),
    name = "Spearman r"
  ) +
  #labs(
  #  title = "Baseline, Stimulation, Oocyte Variables"
 # ) +
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5
    ),
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1,
      size = 8,
      colour = "black"
    ),
    axis.text.y = element_text(
      size = 8,
      colour = "black"
    ),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 9.5),
    plot.margin = margin(5, 5, 5, 5)
  )

corrplot2

ggsave("corrplot2.png", corrplot2, dpi = 600)
```

## Main correlation matrix - atfter reducing some dosing features

```{r, fig.width=10, fig.height=9}
corr_vars3 <- c(
  "Age",
  "BMI",
  "Baseline_AMH",
  "Baseline_total_follicles",
  "Baseline_endometrium",
  "FSH_only",
  #"FSH_initial_dose",
  "hMG_only",
  #"hMG_initial_dose",
  "Dual",
  "Starting_gonadotropin_dose",
  "Total_gonadotropin_dose",
  "GnRH_antagonist_total_dose",
  "hCG_trigger_used",
  "GnRH_agonist_trigger_used",
  "Last_day_of_stim",
  "No_mature_eggs"
)

corr_data3 <- shortdata %>%
  dplyr::select(all_of(corr_vars3))

corr_matrix3 <- cor(
  corr_data3,
  use = "pairwise.complete.obs",
  method = "spearman"
)

corr_matrix3_pretty <- apply_pretty_names(corr_matrix3)

corrplot3 <- corrplot(
  corr_matrix3_pretty,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  addCoef.col = "black",
  number.cex = 0.8,
  tl.cex = 0.9,
  tl.col = "black",
  diag = FALSE
)

corrplot3 <- ggcorrplot(
  corr_matrix3_pretty,
  type = "lower",
  lab = TRUE,
  lab_size = 2.5,
  hc.order = TRUE,
  hc.method = "ward.D2",
  colors = c("#2166AC", "white", "#D6604D"),
  ggtheme = theme_void()
) +
  scale_fill_gradientn(
    colours = c("#2166AC", "white", "#D6604D"),
    limits = c(-1, 1),
    name = "Spearman r"
  ) +
  #labs(
  #  title = "Baseline, Stimulation, Oocyte Variables"
 # ) +
  theme(
    plot.title = element_text(
      size = 14,
      face = "bold",
      hjust = 0.5
    ),
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1,
      size = 11,
      #face = "bold",
      colour = "black"
    ),
    axis.text.y = element_text(
      size = 11,
      #face = "bold",
      colour = "black"
    ),
    legend.title = element_text(size = 9.5),
    legend.text = element_text(size = 10),
    plot.margin = margin(5, 5, 5, 5)
  )

corrplot3

ggsave("corrplot3.png", corrplot3, dpi = 600)
```

```{r, fig.width=12, fig.height=10}
corr_combined <- corrplot1 / corrplot3
corr_combined
```

## Male correlation matrix

```{r}
corr_vars_male <- c(
  "Pre_wash_sperm_conc",
  "Pre_wash_sperm_motility",
  "Pre_wash_normal_sperm_morphology",
  "Post_wash_sperm_conc",
  "Post_wash_sperm_motility",
  "Semen_vol",
  "No_fertilised",
  "Fertilisation_rate"
)

corr_data_male <- shortdata %>%
  dplyr::select(all_of(corr_vars_male))

corr_matrix_male <- cor(
  corr_data_male,
  use = "pairwise.complete.obs",
  method = "spearman"
)

corr_matrix_male_pretty <- apply_pretty_names(corr_matrix_male)

corrplot(
  corr_matrix_male_pretty,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  addCoef.col = "black",
  number.cex = 0.8,
  tl.cex = 0.9,
  tl.col = "black",
  diag = FALSE
)
```

## To make correlation function later

```{r}
make_corr_matrix <- function(data, vars, method = "spearman") {
  
  corr_data <- data %>%
  corr_matrix <- cor(
    corr_data,
    use = "pairwise.complete.obs",
    method = method
  )
  }
  
  return(corr_matrix)
```


```{r}
corr_matrix_fsh <- make_corr_matrix(shortdata, corr_vars_fsh)
corr_matrix_hmg <- make_corr_matrix(shortdata, corr_vars_hmg)
corr_matrix_fsh_hmg <- make_corr_matrix(shortdata, corr_vars_fsh_hmg)
corr_matrix3 <- make_corr_matrix(shortdata, corr_vars3)
```


```{r}
corrplot(
  corr_matrix_fsh,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("#3B6EA8", "white", "#C44E52"))(200),
  addCoef.col = "black",
  number.cex = 0.6,
  tl.cex = 0.8,
  tl.col = "black",
  tl.srt = 45,
  diag = FALSE
)
```

```{r}
corr_data_male <- shortdata %>%
  select(all_of(corr_vars_male))

male_corr_matrix <- cor(
  corr_data_male,
  use = "pairwise.complete.obs",
  method = "spearman"
)

male_corr_matrix_pretty <- apply_pretty_names(male_corr_matrix)

corrplot(
  male_corr_matrix_pretty,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  addCoef.col = "black",
  number.cex = 0.8,
  tl.cex = 0.9,
  tl.col = "black",
  diag = FALSE
)
```
```{r}

## Create dose bins
shortdata <- shortdata %>%
  mutate(dose_bin = cut(
    Starting_gonadotropin_dose,
    breaks = c(-Inf, 225, 350, Inf),
    labels = c("150–225", "226–350", ">350"),
    right = TRUE
  ))


# Compute correlation of MII with each dose bin (dummy-encoded) --> Spearman's 
# Reshape to long format for heatmap
cor_data <- shortdata %>%
  mutate(
    bin_150225 = as.numeric(dose_bin == "150–225"),
    bin_226350 = as.numeric(dose_bin == "226–350"),
    bin_gt350  = as.numeric(dose_bin == ">350")
  ) %>%
  summarise(
    `150–225` = cor(No_mature_eggs, bin_150225, method = "spearman", use = "complete.obs"),
    `226–350` = cor(No_mature_eggs, bin_226350, method = "spearman", use = "complete.obs"),
    `>350`    = cor(No_mature_eggs, bin_gt350,  method = "spearman", use = "complete.obs")
  ) %>%
  tidyr::pivot_longer(everything(), names_to = "dose_bin", values_to = "correlation") %>%
  mutate(variable = "MII")

# Plot
ggplot(cor_data, aes(x = dose_bin, y = variable, fill = correlation)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(correlation, 2)), size = 4, fontface = "bold") +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = 0, limits = c(-1, 1), name = "Spearman r"
  ) +
  scale_x_discrete(limits = c("150–225", "226–350", ">350")) +
  labs(
    title = "Correlation: MII vs Starting Gonadotropin Doses",
    x = "Starting gonadotropin dose (IU)",
    y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text = element_text(size = 11),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 13)
  )
```



## Sensitivity test --> without extreme AMH value (outlier), supplementary 
```{r}
shortdata1 <- shortdata %>%
  filter(Unique_number !="8088599U")

write.csv(
  shortdata1,
  here("data", "processed", "short_sensitivity.csv"),
  row.names = FALSE
)
```
