---
title: "Distributions and Plots"
output: html_document
date: "2026-06-05"
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggcorrplot)
library(patchwork)
library(corrplot)

library(here)
here::here()

df <- read_csv(
  here("data", "processed", "shortprotocol_firstconsentedcycle.csv"),
  name_repair = "minimal"
)
```
## Overall cohort descriptive stats

```{r}
summary_stats <- function(x) {
  x <- x[!is.na(x)]  # remove NAs
  
  data.frame(
    n = length(x),
    mean = mean(x),
    sd = sd(x),
    median = median(x),
    IQR = IQR(x),
    min = min(x),
    max = max(x)
  )
}
```


```{r}
summary_stats1 <- function(x) {
  x <- x[!is.na(x)]
  
  if (length(x) == 0) {
    return(data.frame(
      n = 0,
      mean = NA,
      sd = NA,
      median = NA,
      IQR = NA,
      min = NA,
      max = NA
    ))
  }
  
  data.frame(
    n = length(x),
    mean = mean(x),
    sd = sd(x),
    median = median(x),
    IQR = IQR(x),
    min = min(x),
    max = max(x)
  )
}
```

```{r}
summary_stats1(df$FSH_initial_dose[df$FSH_initial_dose > 0])
1
summary_stats(df$hMG_initial_dose[df$hMG_initial_dose > 0])
```



```{r}
mean(df$Age) #36.1
mean(df$BMI, na.rm = TRUE) #23.1
median(df$Baseline_AMH, na.rm = TRUE) #16.6
median(df$Baseline_endometrium, na.rm = TRUE) #3.8
median(df$Baseline_total_follicles, na.rm = TRUE) #16

mean(df$Total_gonadotropin_dose) ##3061.5
median(df$Total_gonadotropin_dose) #2925
mean(df$Starting_gonadotropin_dose) #298.6
median(df$Starting_gonadotropin_dose) #300

mean(df$GnRH_antagonist_total_dose) #1573.7
median(df$GnRH_antagonist_total_dose) #1500

## DOESNT CONSIDER THE PROCOL, just the gonadotrophin

median(df$FSH_total_dose[df$FSH_total_dose != 0], na.rm = TRUE) #2475
mean(df$FSH_total_dose[df$FSH_total_dose != 0], na.rm = TRUE) #2498.0

median(df$FSH_initial_dose[df$FSH_initial_dose != 0], na.rm = TRUE) #225
mean(df$FSH_initial_dose[df$FSH_initial_dose != 0], na.rm = TRUE) #246.3

median(df$hMG_total_dose[df$hMG_total_dose != 0], na.rm = TRUE) #1650
mean(df$hMG_total_dose[df$hMG_total_dose != 0], na.rm = TRUE) #2046.6

median(df$hMG_initial_dose[df$hMG_initial_dose != 0], na.rm = TRUE) #150
mean(df$hMG_initial_dose[df$hMG_initial_dose != 0], na.rm = TRUE) #207.8

## Should do within protocols

median(df$No_mature_eggs) #9
median(df$No_eggs) #13
```


## MII Outcome Distribution
```{r, fig.width=10, fig.height=5}
## Frequency histogram

median(df$No_mature_eggs) ##9
IQR(df$No_mature_eggs) ##12
mean(df$No_mature_eggs) ##11.42623
sd(df$No_mature_eggs) ##8.241074


p1 <- ggplot(df, aes(x = No_mature_eggs)) +
  geom_histogram(bins = 10, fill = "#4682B4", colour = "white") +
  labs(x = "MII oocyte count",
       y = "Frequency",
       title = "") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 20),   # x and y axis titles
    axis.text  = element_text(size = 20),   # tick labels
    plot.title = element_text(size = 16, face = "bold")
  )

p2 <- ggplot(df, aes(x = log1p(No_mature_eggs))) +
  geom_histogram(bins = 10, fill = "#4682B4", colour = "white") +
  labs(x = "log1p(MII oocyte count)", y = "Frequency") + labs(title = "B") +
  theme_classic()

p2 <- ggplot(df, aes(x = log1p(No_mature_eggs))) +
  geom_histogram(bins = 10, fill = "#4682B4", colour = "white") +
  labs(x = "log1p(MII oocyte count)",
       y = "Frequency",
       title = "") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 20),   # x and y axis titles
    axis.text  = element_text(size = 20),   # tick labels
    plot.title = element_text(size = 16, face = "bold")
  )
  


p3 <- ggplot(df, aes(x = sqrt(No_mature_eggs))) +
  geom_histogram(bins = 10, fill = "#4682B4", colour = "white") +
  labs(x = "sqrt(MII oocyte count)", y = "Frequency") +
  theme_classic()


p4 <- p1 | p2
p4

p1
p2
ggsave("mii_raw_dist.png", p1, dpi = 600)
ggsave("mii_transf_dist.png", p2, dpi = 600)

```

## Baseline Variable Distributions
```{r, fig.width=8, fig.height=8}
plot_hist <- function(data, var, xlab, bins = 10) {
  ggplot(data, aes(x = {{ var }})) +
    geom_histogram(
      bins = bins,
      fill = "#4682B4",
      colour = "white"
    ) +
    theme_classic() +
    labs(
      x = xlab,
      y = "Frequency"
    ) +
    theme(
      axis.title = element_text(size = 20),
      axis.text = element_text(size = 20)
    )
}

p_age <- plot_hist(df, Age, "Age (years)") + labs(title = "")
p_bmi <- plot_hist(df, BMI, "BMI") + labs(title = "")
p_amh <- plot_hist(df, Baseline_AMH, "AMH (pmol/L)") + labs(title = "C")
p_amh_log <- plot_hist(df, log1p(Baseline_AMH), "log1p(AMH) (pmol/L)") + labs(title = "D")
p_follicles <- plot_hist(df, Baseline_total_follicles, "Baseline follicle count") + labs(title = "E")
p_follicles_log <- plot_hist(df, log1p(Baseline_total_follicles), "log1p(Baseline follicle count)") + labs(title = "F")
p_endo <- plot_hist(df, Baseline_endometrium, "Baseline endometrial thickness (mm)") + labs(title = "G")
p_endo_log <- plot_hist(df, log1p(Baseline_endometrium), "log1p(Baseline endometrial thickness (mm))") + labs(title = "H")

baseline_dist_plot <- 
  (p_age | p_bmi) /
  (p_amh | p_amh_log) /
  (p_follicles | p_follicles_log ) /
  (p_endo | p_endo_log ) +
  plot_annotation(
    #title = "Distribution of baseline clinical variables"
  )

ggsave("age_raw_dist.png", p_age, dpi = 600)
ggsave("bmi_raw_dist.png", p_bmi, width = 6, dpi = 600)

baseline_dist_plot
```
## Relationships between baseline and outcome variables
```{r, fig.width=8, fig.height=8}
df <- df %>%
  mutate(
    log_AMH = log1p(Baseline_AMH),
    log_follicles = log1p(Baseline_total_follicles)
  )

plot_vs_mii <- function(data, xvar, xlab) {
  ggplot(data, aes(x = {{ xvar }}, y = No_mature_eggs)) +
    geom_point(alpha = 1, size = 1.5, colour = "#4682B4") +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.7, colour = "#C23B22") +
    theme_classic() +
    labs(
      x = xlab,
      y = "MII oocyte count"
    ) +
    theme(
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 10)
    )
}

p_age_mii  <- plot_vs_mii(df, Age, "Age (years)") + labs(title = "")
p_bmi_mii  <- plot_vs_mii(df, BMI, "BMI") + labs(title = "")
p_amh_mii  <- plot_vs_mii(df, log_AMH, "log1p(AMH) (pmol/L)") + labs(title = "")
p_foll_mii <- plot_vs_mii(df, log_follicles, "log1p(Baseline follicle count)") + labs(title = "")
p_endo_mii <- plot_vs_mii(df, Baseline_endometrium, "Baseline endometrial thickness (mm)") + labs(title = "")

fig_mii_relationships <-
  (p_age_mii | plot_spacer()) /
  (p_amh_mii | plot_spacer()) /
  (p_foll_mii | plot_spacer()) +
  plot_annotation(
    #title = "Relationships between baseline clinical variables and MII oocyte yield"
  )

fig_mii_relationships

ggsave("scatter_baseline_mii.png", fig_mii_relationships, dpi = 600)
```

## Relationships between baseline variables
```{r, fig.width=8, fig.height=3}
plot_pairwise <- function(data, xvar, yvar, xlab, ylab) {
  ggplot(data, aes(x = {{ xvar }}, y = {{ yvar }})) +
    geom_point(alpha = 1, size = 1.5, colour = "#4682B4") +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.9, colour = "#C23B22") +
    theme_classic() +
    labs(
      x = xlab,
      y = ylab
    ) +
    theme(
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10)
    )
}

p_age_amh <- plot_pairwise(
  df,
  Age, log_AMH,
  "Age (years)", "log1p(AMH) (pmol/L)"
) + labs(title = "A")

p_age_foll <- plot_pairwise(
  df,
  Age, log_follicles,
  "Age (years)", "log1p(Baseline follicle count)"
) + labs(title = "B")

p_foll_amh <- plot_pairwise(
  df,
  log_follicles, log_AMH,
  "log1p(Baseline follicle count)", "log1p(AMH) (pmol/L)"
) + labs(title = "C")

fig_predictor_relationships <-
  (p_age_amh | p_age_foll | p_foll_amh) +
  plot_annotation(
    #title = "Relationships between selected baseline predictors"
  )

fig_predictor_relationships
```
## Distributions of summary stimualation variables
```{r}
p_start_dose <- ggplot(df, aes(x = factor(Starting_gonadotropin_dose))) +
  geom_bar(fill = "#4682B4", colour = "white") +
  theme_classic() +
  labs(
    x = "Starting gonadotropin dose (IU)",
    y = "Frequency"
  ) + labs(title = "A")

p_total_dose <- ggplot(df, aes(x = Total_gonadotropin_dose)) +
  geom_histogram(
    binwidth = 500,
    boundary = 0,
    fill = "#4682B4",
    colour = "white"
  ) +
  theme_classic() +
  labs(
    x = "Total gonadotropin dose (IU)",
    y = "Frequency"
  ) + labs(title = "B")

p_antag_dose <- ggplot(df, aes(x = GnRH_antagonist_total_dose)) +
  geom_histogram(
    binwidth = 300,
    boundary = 0,
    fill = "#4682B4",
    colour = "white"
  ) +
  theme_classic() +
  labs(
    x = "Total GnRH antagonist dose",
    y = "Frequency"
  ) + labs(title = "C")

p_stim_duration <- ggplot(df, aes(x = factor(Last_day_of_stim))) +
  geom_bar(fill = "#4682B4", colour = "white") +
  theme_classic() +
  labs(
    x = "Duration of stimulation (days)",
    y = "Frequency"
  ) + labs(title = "D")


treatment_dist_plot <-
  (p_start_dose | p_total_dose) /
  (p_antag_dose | p_stim_duration) +
  plot_annotation(
    #title = "Distribution of stimulation treatment variables"
  )
treatment_dist_plot
```
## FSH and hMG

```{r, fig.width=8, fig.height=5}
p_fsh_start_dose <- df %>%
  filter(FSH_initial_dose > 0) %>%
  ggplot(aes(x = factor(FSH_initial_dose))) +
  geom_bar(fill = "#4682B4", colour = "white") +
  theme_classic() +
  labs(
    title = "A",
    x = "Starting FSH dose (IU)",
    y = "Frequency"
  ) + labs(title = "A")

p_fsh_total_dose <- df %>%
  filter(FSH_total_dose > 0) %>%
  ggplot(aes(x = FSH_total_dose)) +
  geom_histogram(
    binwidth = 500,
    boundary = 0,
    fill = "#4682B4",
    colour = "white"
  ) +
  theme_classic() +
  labs(
    title = "B",
    x = "Total FSH dose (IU)",
    y = "Frequency"
  )


p_hmg_start_dose <- df %>%
  filter(hMG_initial_dose > 0) %>%
  ggplot(aes(x = factor(hMG_initial_dose))) +
  geom_bar(fill = "#4682B4", colour = "white") +
  theme_classic() +
  labs(
    title = "A",
    x = "Starting hMG dose (IU)",
    y = "Frequency"
  ) + labs(title = "C")

p_hmg_total_dose <- df %>%
  filter(hMG_total_dose > 0) %>%
  ggplot(aes(x = hMG_total_dose)) +
  geom_histogram(
    binwidth = 500,
    boundary = 0,
    fill = "#4682B4",
    colour = "white"
  ) +
  theme_classic() +
  labs(
    x = "Total hMG dose (IU)",
    y = "Frequency"
  ) + labs(title = "D")


treatment_dist_plot2 <-
  (p_fsh_start_dose | p_fsh_total_dose) /
  (p_hmg_start_dose | p_hmg_total_dose)
  
treatment_dist_plot2
```

## Dose ~ Baseline variabes

```{r, fig.width=8, fig.height=8}
plot_dose_relationship <- function(data, xvar, yvar, xlab, ylab) {
  ggplot(data, aes(x = {{ xvar }}, y = {{ yvar }})) +
    geom_point(alpha = 1, size = 1.5, colour = "#4682B4") +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.9, colour = "#C23B22") +
    theme_classic() +
    labs(
      x = xlab,
      y = ylab
    ) +
    theme(
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10)
    )
}

p_dose_age <- plot_dose_relationship(
  df,
  Age,
  Total_gonadotropin_dose,
  "Age (years)",
  "Total gonadotropin dose (IU)"
) + labs(title = "A")

p_dose_amh <- plot_dose_relationship(
  df,
  log_AMH,
  Total_gonadotropin_dose,
  "log1p(AMH)",
  "Total gonadotropin dose (IU)"
) + labs(title = "B")

p_dose_foll <- plot_dose_relationship(
  df,
  log_follicles,
  Total_gonadotropin_dose,
  "log1p(Baseline follicle count)",
  "Total gonadotropin dose (IU)"
) + labs(title = "C")

p_dose_age2 <- plot_dose_relationship(
  df,
  Age,
  Starting_gonadotropin_dose,
  "Age (years)",
  "Starting gonadotropin dose (IU)"
) + labs(title = "D")

p_dose_amh2 <- plot_dose_relationship(
  df,
  log_AMH,
  Starting_gonadotropin_dose,
  "log1p(AMH)",
  "Starting gonadotropin dose (IU)"
) + labs(title = "E")

p_dose_foll2 <- plot_dose_relationship(
  df,
  log_follicles,
  Starting_gonadotropin_dose,
  "log1p(Baseline follicle count)",
  "Starting gonadotropin dose (IU)"
) + labs(title = "F")


row1 <- p_dose_age | p_dose_amh | p_dose_foll
row2 <- p_dose_age2 | p_dose_amh2 | p_dose_foll2

dose_baseline_plot <-
  row1 / row2 +
  plot_layout(heights = c(1, 1)) +
  plot_annotation()

dose_baseline_plot
```
## Dose and MII 
```{r}
plot_dose_mii <- function(data, xvar, xlab) {
  ggplot(data, aes(x = {{ xvar }}, y = No_mature_eggs)) +
    geom_jitter(
      width = 0,
      height = 0.25,
      alpha = 0.75,
      size = 2,
      colour = "#4682B4"
    ) +
    geom_smooth(
      method = "lm",
      se = FALSE,
      linewidth = 0.9,
      colour = "#D7301F"
    ) +
    theme_classic() +
    labs(
      x = xlab,
      y = "MII oocyte count"
    ) +
    theme(
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9),
      plot.tag = element_text(face = "bold", size = 13)
    )
}

p_total_gonad_mii <- plot_dose_mii(
  df,
  Total_gonadotropin_dose,
  "Total gonadotropin dose (IU)"
)

p_start_gonad_mii <- plot_dose_mii(
  df,
  Starting_gonadotropin_dose,
  "Starting gonadotropin dose (IU)"
)

dose_mii_plot <-
  p_total_gonad_mii | p_start_gonad_mii +
  plot_annotation(tag_levels = "A")

dose_mii_plot
```
```{r}
p_fsh_total_mii <- df %>%
  filter(!is.na(FSH_total_dose), FSH_total_dose > 0) %>%
  plot_dose_mii(
    FSH_total_dose,
    "Total FSH dose (IU)"
  )

p_hmg_total_mii <- df %>%
  filter(!is.na(hMG_total_dose), hMG_total_dose > 0) %>%
  plot_dose_mii(
    hMG_total_dose,
    "Total hMG dose (IU)"
  )

p_fsh_start_mii <- df %>%
  filter(!is.na(FSH_initial_dose), FSH_initial_dose > 0) %>%
  plot_dose_mii(
    FSH_initial_dose,
    "Starting FSH dose (IU)"
  )

p_hmg_start_mii <- df %>%
  filter(!is.na(hMG_initial_dose), hMG_initial_dose > 0) %>%
  plot_dose_mii(
    hMG_initial_dose,
    "Starting hMG dose (IU)"
  )

dose_mii_plot <-
  (p_fsh_total_mii | p_hmg_total_mii) /
  (p_fsh_start_mii | p_hmg_start_mii) +
  plot_annotation(tag_levels = "A")

dose_mii_plot
```

## Categorical Variables: Aetiology, Stimulation Protocol, Trigger Type

```{r}
p_mii_aetiology <- ggplot(df, aes(x = Aetiology_group, y = Total_gonadotropin_dose)) +
  geom_boxplot(
    width = 0.55,
    fill = "#4682B4",
    colour = "grey30",
    outlier.shape = NA
  ) +
  geom_jitter(
    width = 0.12,
    alpha = 0.65,
    size = 2,
    colour = "grey25"
  ) +
  theme_classic() +
  labs(
    x = "Infertility aetiology group",
    y = "Total Dose"
  ) +
  theme(
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    axis.title = element_text(size = 12)
  )

p_mii_aetiology
```
```{r}
sessionInfo()
packageVersion("here")
packageVersion("tidyverse")
packageVersion("dplyr")
packageVersion("ggplot2")
packageVersion("ggpubr")
packageVersion("ggcorrplot")
packageVersion("patchwork")
```


