---
title: "EDAx"
output: html_document
date: "2026-03-19"
---

```{r}
library(naniar)
library(tidyverse)
library(gridExtra)
library(corrplot)
library(pheatmap)
library(cluster)
library(factoextra)
library(here)
here::here()
```

```{r}
freshdata <- read_csv(
  here("data","processed","freshdata_cleaned.csv"),
  name_repair = "minimal"
)

freshdata$Unique_number <- NULL
##for now
freshdata$Baseline_follicle_size <- NULL
freshdata$Second_scan_follicle_size <- NULL
freshdata$Final_follicle_size <- NULL

## Define outcome variables
outcomes <- c("No_mature_eggs", "No_fertilised", "Fertilisation_rate")

predictors <- setdiff(names(freshdata), c(outcomes))

## separate continous and categorical predictors
categorical_vars <- c("Aetiology", "PGD", "Male_factor", "Funding", "Treatment", "Protocol_type", "Sperm_source", "Sperm_prep", "Sperm_mobility")
continuous_vars <- setdiff(predictors, categorical_vars)
```

## MISSINGNESS
```{r}
# Table: percentage missing per variable, sorted descending
miss_table <- freshdata %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Pct_missing") %>%
  arrange(desc(Pct_missing)) %>%
  filter(Pct_missing > 0)
 
print(miss_table, n = 50)
 
# Flag variables with >20% missingness — candidates for removal
high_miss <- miss_table %>% filter(Pct_missing > 20)
print(high_miss)
```
```{r}
# Plot 1: missingness bar chart
ggplot(miss_table, aes(x = reorder(Variable, Pct_missing), y = Pct_missing)) +
  geom_col(fill = "#5DCAA5") +
  coord_flip() +
  labs(title = "Percentage missing per variable",
       x = NULL, y = "% missing") +
  theme_minimal()
```
```{r}
# Plot 2: missingness heatmap — shows co-missingness patterns
vis_miss(freshdata, sort_miss = TRUE) +
  labs(title = "Missingness heatmap") +
  theme(axis.text.x = element_text(angle = 90, size = 7))
```
## Descriptive Statistics
```{r}
desc_stats <- describe(freshdata[, continuous_vars])
print(desc_stats[, c("n","mean","sd","median","mad","min","max","skew","kurtosis")])
```
## Distributions
```{r}
plot_distributions <- function(vars, data, batch_size = 12) {
  batches <- split(vars, ceiling(seq_along(vars) / batch_size))
  for (i in seq_along(batches)) {
    plots <- lapply(batches[[i]], function(v) {
      ggplot(data, aes(x = .data[[v]])) +
        geom_histogram(aes(y = after_stat(density)), bins = 30,
                       fill = "#9FE1CB", colour = "white", alpha = 0.8) +
        geom_density(colour = "#0F6E56", linewidth = 0.8) +
        labs(title = v, x = NULL, y = NULL) +
        theme_minimal(base_size = 9)
    })
    do.call(gridExtra::grid.arrange, c(plots, ncol = 3))
  }
}
 
plot_distributions(continuous_vars, freshdata)
```

```{r}
# Outcome variable distributions
outcome_plots <- lapply(outcomes, function(v) {
  ggplot(freshdata, aes(x = .data[[v]])) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   fill = "#AFA9EC", colour = "white") +
    geom_density(colour = "#3C3489", linewidth = 1) +
    labs(title = paste("Distribution:", v)) +
    theme_minimal()
})

do.call(gridExtra::grid.arrange, c(outcome_plots, ncol = 2))
```
## Correlation
```{r}
#Big heatmap of all variables
cont_data <- freshdata[, continuous_vars] %>%
  select(where(~ sum(!is.na(.)) > 10))  # drop near-empty columns

#Remove columns with zero variance first
cont_data_clean <- cont_data %>%
  select(where(~ var(., na.rm = TRUE) > 0)) %>%   # drop zero variance
  select(where(~ sum(!is.na(.)) > 10))              # drop near-empty
#Compute correlation
cor_matrix <- cor(cont_data_clean, method = "spearman", use = "pairwise.complete.obs")

# Check for any remaining NA/NaN in the matrix
cat("Any NAs in correlation matrix:", anyNA(cor_matrix), "\n")

# Remove any rows/cols that still have NA correlations
na_cols <- apply(cor_matrix, 2, function(x) any(is.na(x)))
cat("Columns with NA correlations:", names(na_cols)[na_cols], "\n")
cor_matrix <- cor_matrix[!na_cols, !na_cols]

# Now plot safely
corrplot(cor_matrix,
         method = "color",
         type = "upper",           # just the upper triangle
         order = "hclust",
         hclust.method = "ward.D2",
         tl.cex = 0.4,
         tl.col = "black",
         diag = FALSE,             # hide the diagonal (self-correlations = 1)
         col = colorRampPalette(c("#185FA5","white","#993C1D"))(200),
         title = "Spearman correlation — all continuous predictors (clustered)",
         mar = c(0,0,2,0))

```
## Flag highly correlated pairs
```{r}
cor_df <- as.data.frame(as.table(cor_matrix)) %>%
  rename(Var1 = Var1, Var2 = Var2, r = Freq) %>%
  filter(Var1 != Var2, abs(r) > 0.7) %>%
  mutate(pair = paste(pmin(as.character(Var1), as.character(Var2)),
                      pmax(as.character(Var1), as.character(Var2)), sep = " ~ ")) %>%
  distinct(pair, .keep_all = TRUE) %>%
  arrange(desc(abs(r)))

print(cor_df[, c("pair", "r")], n = 40)

###this helps with parameter reduction
```

## Each predictor vs. primary outcome: number of mature eggs
```{r}
#Remove zero variance variables first
valid_vars <- continuous_vars[sapply(continuous_vars, function(v) {
  x <- suppressWarnings(as.numeric(as.character(freshdata[[v]])))
  sd(x, na.rm = TRUE) > 0 && sum(!is.na(x)) > 10
})]

print(continuous_vars[!continuous_vars %in% valid_vars])

#Now run correlation safely
outcome_cor <- sapply(valid_vars, function(v) {
  x <- suppressWarnings(as.numeric(as.character(freshdata[[v]])))
  cor(x, freshdata$No_mature_eggs, method = "spearman", use = "complete.obs")
})

outcome_cor_df <- data.frame(
  Variable = names(outcome_cor),
  r = outcome_cor
) %>%
  filter(!is.na(r)) %>%
  arrange(desc(abs(r)))

ggplot(outcome_cor_df, aes(x = reorder(Variable, r), y = r,
                            fill = r > 0)) +
  geom_col() +
  scale_fill_manual(values = c("#D85A30","#1D9E75"), guide = "none") +
  coord_flip() +
  labs(title = "Spearman correlation with Number of mature eggs",
       x = NULL, y = "Spearman r") +
  theme_minimal(base_size = 9)
```
```{r}
#Top correlated predictors with No_mature_eggs
top_correlations <- outcome_cor_df %>%
  arrange(desc(abs(r))) %>%
  mutate(
    r = round(r, 3),
    abs_r = round(abs(r), 3),
    direction = ifelse(r > 0, "positive", "negative")
  )

#Print top 20
print(top_correlations, n = 20)

#Or set your own threshold — e.g. only show |r| > 0.3
#strong_correlations <- top_correlations %>% filter(abs_r > 0.3)
#print(strong_correlations)

#Save to CSV for your report
#write.csv(top_correlations, "outcome_correlations.csv", row.names = FALSE)
```

## Each predictor vs. primary outcome: number of mature eggs
```{r}
#Now run correlation safely
outcome_cor1 <- sapply(valid_vars, function(v) {
  x <- suppressWarnings(as.numeric(as.character(freshdata[[v]])))
  cor(x, freshdata$Fertilisation_rate, method = "spearman", use = "complete.obs")
})

outcome_cor1_df <- data.frame(
  Variable = names(outcome_cor1),
  r = outcome_cor1
) %>%
  filter(!is.na(r)) %>%
  arrange(desc(abs(r)))

ggplot(outcome_cor1_df, aes(x = reorder(Variable, r), y = r,
                            fill = r > 0)) +
  geom_col() +
  scale_fill_manual(values = c("#D85A30","#1D9E75"), guide = "none") +
  coord_flip() +
  labs(title = "Spearman correlation with Fertilisation Rate",
       x = NULL, y = "Spearman r") +
  theme_minimal(base_size = 9)
```
```{r}
#Top correlated predictors with Fertilisation_rate
top_correlations1 <- outcome_cor1_df %>%
  arrange(desc(abs(r))) %>%
  mutate(
    r = round(r, 3),
    abs_r = round(abs(r), 3),
    direction = ifelse(r > 0, "positive", "negative")
  )

#Print top 20
print(top_correlations1, n = 20)
```

## Stats

```{r}
freshdata$Protocol_type <- as.factor(freshdata$Protocol_type)
freshdata$Treatment <- as.factor(freshdata$Treatment)
freshdata$Aetiology <- as.factor(freshdata$Aetiology)

print(levels(freshdata$Protocol_type))
print(levels(freshdata$Treatment))
print(levels(freshdata$Aetiology))
```
## Test whether fertilisation rates differs across categorical groups

```{r}
kw_protocol <- kruskal.test(Fertilisation_rate ~ Protocol_type, data = freshdata)
print(kw_protocol)
```
```{r}
kw_treatment <- kruskal.test(Fertilisation_rate ~ Treatment, data = freshdata)
print(kw_treatment)
```
```{r}
kw_aetiology <- kruskal.test(Fertilisation_rate ~ Aetiology, data = freshdata)
print(kw_aetiology)
```
## Test whether number of mature oocytes differs between categorical variables
```{r}
kw_protocol1 <- kruskal.test(No_mature_eggs ~ Protocol_type, data = freshdata)
print(kw_protocol1)
```
```{r}
kw_treatment1 <- kruskal.test(No_mature_eggs ~ Treatment, data = freshdata)
print(kw_treatment1)
```
```{r}
kw_aetiology1 <- kruskal.test(No_mature_eggs ~ Aetiology, data = freshdata)
print(kw_aetiology1)
```
## Clustering

```{r}
cluster_vars <- c("Age", "Baseline_total_follicles", "FSH_total_dose", "FSH_duration", "Last_day_of_stim", "Final_right_follicles","Final_left_follicles", "No_mature_eggs", "Fertilisation_rate")


# Extract and clean clustering variables
cluster_data_raw <- freshdata[, intersect(cluster_vars, names(freshdata))] %>%
  mutate(across(everything(), 
                ~ suppressWarnings(as.numeric(as.character(.)))))

#Keep track of which rows are complete (needed to map clusters back)
complete_rows <- which(complete.cases(cluster_data_raw))
cat("Patients with complete clustering variables:", length(complete_rows), "\n")
cat("Patients dropped due to missing values:", nrow(freshdata) - length(complete_rows), "\n")

```

```{r}
# Scale the data — essential before clustering
cluster_scaled <- cluster_data_raw[complete_rows, ] %>% scale()
```

```{r}
wss <- sapply(1:8, function(k) {
  pam(cluster_scaled, k)$objective[1]
})

##Elbow plot suggests 3 clusters is optimal
plot(1:8, wss, type = "b", pch = 19, col = "#1D9E75",
     xlab = "Number of clusters (k)",
     ylab = "Total within-cluster distance",
     main = "Elbow method — optimal k")
```
```{r}
# Silhouette widths across k values
avg_sil <- sapply(2:8, function(k) {
  p <- pam(cluster_scaled, k)
  s <- silhouette(p$clustering, dist(cluster_scaled))
  mean(s[, 3])
})
plot(2:8, avg_sil, type = "b", pch = 19, col = "#7F77DD",
     xlab = "Number of clusters (k)",
     ylab = "Average silhouette width",
     main = "Silhouette method — optimal k")
abline(v = which.max(avg_sil) + 1, lty = 2, col = "red")
cat("Optimal k by silhouette:", which.max(avg_sil) + 1, "\n")

##3 clusters optimal
```

```{r}
#Set k and run PAM
k <- 3
set.seed(42)
pam_result <- pam(cluster_scaled, k)

cat("\nCluster sizes:\n")
print(table(pam_result$clustering))
```

```{r}
sil <- silhouette(pam_result$clustering, dist(cluster_scaled))
plot(sil,
     col = c("#1D9E75","#7F77DD","#D85A30"),
     border = NA,
     main = "Silhouette plot — k=3")
cat("Average silhouette width:", round(mean(sil[, 3]), 3), "\n")
```

```{r}
clusplot(cluster_scaled,
         pam_result$clustering,
         color = TRUE,
         col.p = c("#1D9E75","#7F77DD","#D85A30")[pam_result$clustering],
         col.clus = c("#1D9E75","#7F77DD","#D85A30"),
         shade = FALSE,
         lines = 0,
         main = "PAM clustering — k=3",
         sub = "")
```

## Map clusters back to original data
```{r}
freshdata$cluster <- NA
freshdata$cluster[complete_rows] <- pam_result$clustering
freshdata$cluster <- as.factor(freshdata$cluster)

#Check it worked
cat("Cluster assignments:\n")
print(table(freshdata$cluster, useNA = "always"))
```
## Cluster stats
```{r}
cluster_summary <- freshdata %>%
  filter(!is.na(cluster)) %>%
  group_by(cluster) %>%
  summarise(
    n                  = n(),
    median_mature_eggs = round(median(No_mature_eggs,     na.rm = TRUE), 1),
    IQR_mature_eggs    = round(IQR(No_mature_eggs,        na.rm = TRUE), 1),
    median_fert_rate   = round(median(Fertilisation_rate, na.rm = TRUE), 3),
    median_age         = round(median(Age,                na.rm = TRUE), 1),
    median_total_fol         = round(median(Baseline_total_follicles,       na.rm = TRUE), 2),
    median_FSH_dose    = round(median(FSH_total_dose,     na.rm = TRUE), 0),
    median_FSH_duration = round(median(FSH_duration, na.rm = TRUE), 0),
    median_stim_days   = round(median(Last_day_of_stim,   na.rm = TRUE), 1)
)
print(cluster_summary)
#write.csv(cluster_summary, "cluster_characterisation.csv", row.names = FALSE)

#Age", "Baseline_total_follicles", "FSH_total_dose", "FSH_duration", "Last_day_of_stim", "Final_right_follicles","Final_left_follicles", "No_mature_eggs", "Fertilisation_rate
```

```{r}
# Mature oocytes by cluster
ggplot(freshdata %>% filter(!is.na(cluster)),
       aes(x = cluster, y = No_mature_eggs, fill = cluster)) +
  geom_boxplot(outlier.shape = 16, outlier.alpha = 0.5, alpha = 0.8) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
  scale_fill_manual(values = c("#9FE1CB","#CECBF6","#F5C4B3"), guide = "none") +
  labs(title = "Mature oocytes by cluster",
       x = "Cluster", y = "No. mature oocytes") +
  theme_minimal()
```

```{r}
# Fertilisation rate by cluster  
ggplot(freshdata %>% filter(!is.na(cluster)),
       aes(x = cluster, y = Fertilisation_rate, fill = cluster)) +
  geom_boxplot(outlier.shape = 16, outlier.alpha = 0.5, alpha = 0.8) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
  scale_fill_manual(values = c("#9FE1CB","#CECBF6","#F5C4B3"), guide = "none") +
  labs(title = "Fertilisation rate by cluster",
       x = "Cluster", y = "Fertilisation rate") +
  theme_minimal()
```



```{r}

# AMH by cluster — shows whether clusters reflect ovarian reserve
#ggplot(freshdata %>% filter(!is.na(cluster)),
     #  aes(x = cluster, y = Baseline_AMH, fill = cluster)) +
#  geom_boxplot(outlier.shape = 16, outlier.alpha = 0.5, alpha = 0.8) +
 # geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
 # scale_fill_manual(values = c("#9FE1CB","#CECBF6","#F5C4B3"), guide = "none") +
 # labs(title = "Baseline AMH by cluster",
  #     x = "Cluster", y = "AMH") +
 # theme_minimal()

##8

#kw_eggs <- kruskal.test(No_mature_eggs ~ cluster,
 #                        data = freshdata %>% filter(!is.na(cluster)))
#kw_fert <- kruskal.test(Fertilisation_rate ~ cluster,
 #                        data = freshdata %>% filter(!is.na(cluster)))
#
#cat("\nKruskal-Wallis: mature oocytes across clusters\n")
#print(kw_eggs)
#cat("\nKruskal-Wallis: fertilisation rate across clusters\n")
#print(kw_fert)

# Post-hoc if significant
#if (kw_eggs$p.value < 0.05) {
 # cat("\nDunn post-hoc — mature oocytes:\n")
  #dunn.test(freshdata$No_mature_eggs[!is.na(freshdata$cluster)],
   #         g = freshdata$cluster[!is.na(freshdata$cluster)],
    #        method = "bonferroni", label = TRUE)

```

```{r}
dist_matrix <- dist(cluster_scaled, method = "euclidean")
hclust_result <- hclust(dist_matrix, method = "ward.D2")

plot(hclust_result,
     main = "Hierarchical clustering dendrogram",
     xlab = "Patients",
     ylab = "Height",
     cex = 0.4,    # label size — small because many patients
     hang = -1,    # hang labels at the bottom
     labels = FALSE)  # hide individual patient labels — too many to read

# Add coloured rectangles to show k=3 clusters
rect.hclust(hclust_result, k = 3, 
            border = c("#1D9E75","#7F77DD","#D85A30"))

```
```{r}
k <- 3  # use same k as PAM for comparison
hclust_labels <- cutree(hclust_result, k = k)

cat("\nHierarchical cluster sizes:\n")
print(table(hclust_labels))
```
```{r}
sil_hclust <- silhouette(hclust_labels, dist_matrix)
plot(sil_hclust,
     col = c("#1D9E75","#7F77DD","#D85A30"),
     border = NA,
     main = "Silhouette plot — hierarchical clustering k=3")

cat("Average silhouette width:", round(mean(sil_hclust[, 3]), 3), "\n")

```

```{r}
clusplot(cluster_scaled,
         hclust_labels,
         color = TRUE,
         col.p = c("#1D9E75","#7F77DD","#D85A30")[hclust_labels],
         col.clus = c("#1D9E75","#7F77DD","#D85A30"),
         shade = FALSE,
         lines = 0,
         main = "Hierarchical clustering — k=3",
         sub = "")
```
```{r}
freshdata$hclust_cluster <- NA
freshdata$hclust_cluster[complete_rows] <- hclust_labels
freshdata$hclust_cluster <- as.factor(freshdata$hclust_cluster)
```

```{r}
cat("\nPAM vs hierarchical cluster agreement:\n")
print(table(PAM = freshdata$cluster, 
            Hierarchical = freshdata$hclust_cluster,
            useNA = "no"))

# Agreement rate
agree <- mean(as.integer(freshdata$cluster) == 
              as.integer(freshdata$hclust_cluster),
              na.rm = TRUE)
cat("Raw agreement rate:", round(agree * 100, 1), "%\n")
cat("Note: cluster numbers may not match between methods\n")
cat("Look at the cross-tab to see if clusters correspond\n") 
```

