---
title: "Exploratory Regression Modelling for Thesis"
output: html_document
date: "2026-06-03"
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(MASS)
library(patchwork)
library(broom)
library(car)
library(forcats)
library(here)
here::here()

df <- read_csv(
  here("data","processed","shortprotocol_firstconsentedcycle.csv"),
  name_repair = "minimal",
)
```

## Factor categorical variables and log-trasnform skewed
```{r}
df$Aetiology_group <- factor(df$Aetiology_group)
df$Aetiology_group <- relevel(
  df$Aetiology_group,
  ref = "Female_factor"
)

df$Treatment_type <- factor(df$Treatment_type)
df$Treatment_type <- relevel(
  df$Treatment_type,
  ref = "FSH_only"
)

df <- df %>%
  mutate(
    log_MII = log1p(No_mature_eggs),
    log_AMH = log1p(Baseline_AMH),
    log_baseline_follicles = log1p(Baseline_total_follicles)
  )
```

## GLM not suited to data - heteroscedascitiy 
```{r}
glm <- lm(No_mature_eggs ~ Age + BMI + log_AMH + log_baseline_follicles + Baseline_endometrium + Aetiology_group,
  data = df)

plot(glm)
```

## Fit Poisson model first and check whether residual variation is larger than expected
```{r}
poisson_model <- glm(
  No_mature_eggs ~ Age + BMI + log_AMH + log_baseline_follicles +
    Baseline_endometrium + Aetiology_group,
  family = poisson(link = "log"),
  data = df
)

overdispersion_ratio <- sum(residuals(poisson_model, type = "pearson")^2) /
  poisson_model$df.residual

overdispersion_ratio #2.13589
```

## NEGATIVE BINOMIAL REGRESSION

## Model 1:  pre-treatment baseline
```{r}
model_nb1 <- glm.nb(
  No_mature_eggs ~ Age + BMI + log_AMH + log_baseline_follicles + Baseline_endometrium + Aetiology_group,
  data = df
)

summary(model_nb1)
exp(coef(model_nb1))


nb_table <- tidy(model_nb1, conf.int = TRUE, exponentiate = TRUE) %>%
  mutate(
    estimate = round(estimate, 2),
    conf.low = round(conf.low, 2),
    conf.high = round(conf.high, 2),
    p.value = signif(p.value, 3)
  )

nb_table
```

## Model 2: pre-treatment baseline + regimen
```{r}
model_nb2 <- glm.nb(
  No_mature_eggs ~ Age + BMI + log_baseline_follicles + log_AMH + Treatment_type + Aetiology_group,
  data = df
)

summary(model_nb2)
exp(coef(model_nb2))
```

## Model 3: pre-treatment baseline + starting dose
```{r}
model_nb3 <- glm.nb(
  No_mature_eggs ~ Age + BMI + log_AMH * Starting_gonadotropin_dose + log_baseline_follicles + Baseline_endometrium + Aetiology_group,
  data = df
)

summary(model_nb3)
exp(coef(model_nb3))
```

## Model 3: centre AMH and starting dose
```{r}
df <- df %>%
  mutate(
    log_AMH_c = log_AMH - mean(log_AMH, na.rm = TRUE),
    starting_dose_c =
      Starting_gonadotropin_dose -
      mean(Starting_gonadotropin_dose, na.rm = TRUE)
  )

df <- df %>%
  mutate(
    log_AMH_c = log_AMH - mean(log_AMH, na.rm = TRUE),

    starting_dose_50_c = (
      Starting_gonadotropin_dose -
      mean(Starting_gonadotropin_dose, na.rm = TRUE)
    ) / 50
  )

nb4 <- glm.nb(
  No_mature_eggs ~ Age + BMI + log_baseline_follicles + Baseline_endometrium + Aetiology_group + log_AMH_c * starting_dose_50_c, data = df
)

summary(nb4)
exp(coef(nb4))
```

## Make figure plots:

```{r}
predictor_labels <- c(
  "Age"                                   = "Age",
  "BMI"                                   = "BMI",
  "log_AMH"                               = "AMH",
  "Baseline_AMH"                          = "AMH",
  "log_baseline_follicles"                = "Total follicle count",
  "Baseline_total_follicles"              = "Total follicle count",
  "Baseline_endometrium"                  = "Endometrial thickness",
  "Aetiology_groupNo_female_factor"       = "No female factor",
  "Aetiology_groupUnexplained"            = "Unexplained infertility",
  "Treatment_typeDual"                    = "Dual stimulation",
  "Treatment_typehMG_only"                = "hMG only",
  "Starting_gonadotropin_dose"            = "Starting gonadotropin dose",
  "Total_gonadotropin_dose"               = "Total gonadotropin dose",
  "Stimulation_duration"                  = "Stimulation duration",
  "log_AMH:Starting_gonadotropin_dose"    = "AMH × starting dose",
  "log_AMH_c" = "AMH",
  "starting_dose_50_c" = "Starting dose (per 50 IU)",
  "log_AMH_c:starting_dose_50_c" = "AMH × starting dose"
)

recode_predictors <- function(x) {
  
  new_labels <- unname(predictor_labels[x])
  
  # Keep the original term when no replacement is specified
  new_labels[is.na(new_labels)] <- x[is.na(new_labels)]
  
  new_labels
}
```

```{r}
prepare_irr_data <- function(model, labels = predictor_labels) {

  broom::tidy(model, conf.int = TRUE) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      IRR = exp(estimate),
      conf.low.IRR = exp(conf.low),
      conf.high.IRR = exp(conf.high),

      Predictor = dplyr::recode(
        term,
        !!!labels,
        .default = term
      ),

      Direction = case_when(
        IRR > 1 ~ "Higher expected MII count",
        IRR < 1 ~ "Lower expected MII count",
        TRUE ~ "No association"
      ),

      Significant = p.value < 0.05
    )
}
```

```{r}
plot_irr_direction <- function(
    model,
    title = NULL,
    labels = predictor_labels,
    #x_breaks = c(0.5, 0.75, 1, 1.5, 2, 3),
    predictor_order = NULL
) {

  plot_data <- prepare_irr_data(
    model = model,
    labels = labels
  )

  # Either use a specified common order or order by IRR
  if (is.null(predictor_order)) {

    plot_data <- plot_data %>%
      mutate(
        Predictor = fct_reorder(Predictor, IRR)
      )

  } else {

    plot_data <- plot_data %>%
      mutate(
        Predictor = factor(
          Predictor,
          levels = rev(predictor_order)
        )
      )
  }

  ggplot(
    plot_data,
    aes(x = IRR, y = Predictor)
  ) +

    # IRR = 1 represents no association
    geom_vline(
      xintercept = 1,
      linetype = "dashed",
      colour = "grey40",
      linewidth = 0.8
    ) +

    # Connect each estimate to the null value
    geom_segment(
      aes(
        x = 1,
        xend = IRR,
        yend = Predictor,
        colour = Direction
      ),
      linewidth = 1.1
    ) +

    # Filled = significant; open = non-significant
    geom_point(
      aes(
        colour = Direction,
        shape = Significant
      ),
      size = 2.5,
      stroke = 1.2
    ) +

    scale_colour_manual(
      values = c(
        "Lower expected MII count" = "#2166AC",
        "Higher expected MII count" = "red",
        "No association" = "grey50"
      )
    ) +

    scale_shape_manual(
      values = c(
        "TRUE" = 16,
        "FALSE" = 1
      ),
      breaks = c(TRUE, FALSE),
      labels = c(
        "p < 0.05",
        "p ≥ 0.05"
      )
    ) +

    #scale_x_log10(
    #  breaks = x_breaks,
    #  labels = format(x_breaks, trim = TRUE),
    #  expand = expansion(mult = c(0.08, 0.08))
   # ) +
    
    #scale_x_continuous(
    # breaks = c(1.0, 1.25, 1.5, 1.75, 2.0),
    # labels = c("1.00", "1.25", "1.50", "1.75", "2.00"),
     #limits = c(0.88, 2.20),
    # expand = expansion(mult = c(0, 0))
    # ) +
    scale_x_continuous(
      breaks = scales::breaks_pretty(n = 7),
      labels = scales::label_number(accuracy = 0.01),
      expand = expansion(mult = c(0.05, 0.08))
      ) +
    
    labs(
      title = title,
      x = "Incidence rate ratio (IRR)",
      y = NULL,
      colour = NULL,
      shape = "Statistical significance"
    ) +

    theme_classic(base_size = 12) +

    theme(
      plot.title = element_text(
        face = "bold",
        size = 15,
        hjust = 0.5,
        margin = margin(b = 10)
      ),

      axis.title.x = element_text(
        #face = "bold",
        size = 16,
        margin = margin(t = 10)
      ),

      axis.text.x = element_text(
        face = "bold",
        size = 14,
      ),

      axis.text.y = element_text(
        #face = "bold",
        size = 16,
        margin = margin(r = 8)
      ),

      axis.line = element_line(linewidth = 0.7),
      axis.ticks = element_line(linewidth = 0.7),

      legend.position = "bottom",
      legend.text = element_text(size = 14),
      legend.title = element_text(
        #face = "bold",
        size = 14
      ),

      plot.margin = margin(10, 15, 10, 10)
    )
}
```

```{r}
p_nb1 <- plot_irr_direction(
  model = model_nb1,
  #title = "Baseline model"
)

p_nb1

ggsave("../figures/nb1.png", p_nb1, width = 12, height = 6,  dpi = 600)
```

```{r, fig.width=12, fig.height=5}
p_nb2 <- plot_irr_direction(
  model = model_nb2,
  #title = "Stimulation regimen model"
)

p_nb2

ggsave("../figures/nb2.png", p_nb2, width = 12, height = 6, dpi = 600)
```

```{r}
p_nb4 <- plot_irr_direction(
  model = nb4
)

p_nb4

ggsave("nb44.png", p_nb4, width = 15, height = 6, dpi = 600)
```



## Fit Poisson Generalized Linear Model and see overdispersion, if overdispersion is present --> switch to negative binomial regression
## Compare poisson and nbd residuals plots --> they look the samme --> supplementary figure?
```{r}

# Poisson
plot(
  fitted(poisson_model), rstandard(poisson_model),
  col = "steelblue",
  pch = 16,
  xlab = "Predicted MII",
  ylab = "Standardised residuals",
  main = "Poisson"
)
abline(h = 0, lty = 2)
abline(h = c(-2, 2), lty = 3, col = "grey40")

# Negative binomial
plot(
  fitted(model1), rstandard(model1),
  col = "steelblue",
  pch = 16,
  xlab = "Predicted MII",
  ylab = "Standardised residuals",
  main = "Negative binomial"
)
abline(h = 0, lty = 2)
abline(h = c(-2, 2), lty = 3, col = "grey40")

par(mfrow = c(1, 1))
```

```{r}
## Poisson GLM
poisson_model <- glm(
  No_mature_eggs ~ Age + BMI + loBaseline_AMH + Baseline_total_follicles + Baseline_endometrium,
  data = df,
  family = poisson(link = "log")
)

## Overdispersion checks ---> Over dispersion is present
dispersion_ratio <- deviance(poisson_model) / df.residual(poisson_model)
dispersion_ratio ##3.571943

pearson_dispersion <- sum(residuals(poisson_model, type = "pearson")^2) /
  df.residual(poisson_model) ##3.261497
pearson_dispersion

## Negative Binomial Model
nb1 <- glm.nb(
  No_mature_eggs ~ Age + BMI + log_AMH + log_baseline_follicles + Baseline_endometrium + Aetiology_group,
  data = df
)

AIC(poisson_model, nb1) ## lower AIC --> negative binomial better suited to data

summary(nb1)
exp(coef(nb1))

tidy(nb1, exponentiate = TRUE, conf.int = TRUE)
```



