---
title: "Linear and Negative Binomial Modelling"
output: html_document
date: "2026-04-08"
---

## How well can an outcome be predicted before treatment?

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)

library(naniar)
library(gridExtra)
library(corrplot)
library(pheatmap)
library(cluster)
library(factoextra)
library(Hmisc)
library(mice)
library(here)
library(lmtest)
library(car)
library(MASS)
here::here()

## ALL DATA
freshdata <- read_csv(
  here("data","processed","freshdata_cleaned1.csv"),
  name_repair = "minimal"
)

## SHORT ANTAGONIST + FERTILITY PRESERVATION, NO EGG DONOR, N=53
shortdata <- freshdata %>%
  dplyr::filter(Protocol_type == "Short Antagonist")
##add line except for egg donor
```

```{r}
short_data <- freshdata %>%
  dplyr::filter(Protocol_type == "Short Antagonist") %>%
  dplyr::select(
    Unique_number,
    No_mature_eggs,
    Age,
    BMI,
    Baseline_AMH,
    Baseline_total_follicles,
    Baseline_endometrium,
    Aetiology_group,
    FSH_used,
    hMG_used,
    GnRH_antagonist_duration,
    GnRH_antagonist_total_dose,
    FSH_total_dose,
    FSH_duration,
    FSH_initial_dose,
    FSH_final_dose,
    FSH_min_dose,
    FSH_max_dose,
    FSH_dose_changes,
    FSH_avg_daily,
    hMG_total_dose,
    hMG_duration,
    hMG_initial_dose,
    hMG_final_dose,
    hMG_max_dose,
    hMG_min_dose,
    hMG_avg_daily,
    Total_stim_dose,
    hCG_trigger_used,
    hCG_trigger_dose,
    GnRH_trigger_used,
    GnRH_agonist_trigger_dose,
    Last_day_of_stim,
    Final_follicles_lessthan_11.9
  )

short_data1 <- freshdata %>%
  dplyr::filter(Protocol_type == "Short Antagonist") %>%
  dplyr::select(
    No_mature_eggs,
    Age,
    BMI,
    Baseline_AMH,
    Baseline_total_follicles,
    Baseline_endometrium,
    Aetiology_group
  ) %>%
  mutate(
    Aetiology_group = as.factor(Aetiology_group)
  )

colSums(is.na(short_data1))
md.pattern(short_data1)

summary(freshdata$Baseline_AMH)
```

## Impute missingness with MICE
```{r}
meth <- make.method(shortdata)
pred <- make.predictorMatrix(shortdata)

# Do not impute the outcome
meth["No_mature_eggs"] <- ""

# Keep outcome available as a predictor for imputing covariates
pred[, "No_mature_eggs"] <- 1
pred["No_mature_eggs", ] <- 0

# Methods
meth["BMI"] <- "pmm"
meth["Baseline_AMH"] <- "pmm"
meth["Baseline_total_follicles"] <- "pmm"
meth["Baseline_endometrium"] <- "pmm"

imp <- mice(
  shortdata,
  m = 20,
  method = meth,
  predictorMatrix = pred,
  seed = 123,
  printFlag = FALSE
)
```

## Data Checks
```{r}
hist(comp1$No_mature_eggs, breaks = 15, main = "Raw mature oocytes", xlab = "No_mature_eggs")
hist(log1p(comp1$No_mature_eggs), breaks = 15, main = "log1p mature oocytes", xlab = "log1p(No_mature_eggs)")
table(comp1$Aetiology_group, useNA = "ifany")
```

##Fit linear model
```{r}
##normal mature oocytes
lm1 <- with(
  imp,
  lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles +
       Baseline_endometrium + Aetiology_group)
)

summary(pool(lm1))
```

```{r}
##log mature oocytes
lm2 <- with(
  imp,
  lm(log1p(No_mature_eggs) ~ Age + BMI + Baseline_AMH + Baseline_total_follicles +
       Baseline_endometrium + Aetiology_group)
)

summary(pool(lm2))
```

## Check LM assumptions with MICE of non-log(y) model
```{r}
#Extract 1 completed dataset, fit the same model on the completed dataset, inspect assumptions, can repeat on 2-3 datasets
comp1 <- complete(imp, 1)

lm1.1 <- lm(
  No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles +
    Baseline_endometrium + Aetiology_group,
  data = comp1
)

summary(lm1.1)

plot(lm1.1)
shapiro.test(residuals(lm1.1)) #p=0.4841
bptest(lm1.1) ##0.1812
vif(lm1.1)
```
## Check LM assumptions with MICE of log(y) model
```{r}
lm2.1 <- lm(
  log1p(No_mature_eggs) ~ Age + BMI + Baseline_AMH + Baseline_total_follicles +
    Baseline_endometrium + Aetiology_group,
  data = comp1
)

summary(lm2.1)

plot(lm2.1)

shapiro.test(residuals(lm2.1)) ##p=0.3475
bptest(lm2.1) #0.1814
vif(lm2.1)
```
##Fit candidate models, inspect pooled coefficients and significance, fit the same candidate models on comp1, compare values

## CORE baseline
```{r}
fit_lm1 <- with(
  imp,
  lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles)
)
summary(pool(fit_lm1))
```
## Add endometrium
```{r}
fit_lm2 <- with(
  imp,
  lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles + Baseline_endometrium)
)
summary(pool(fit_lm2))
```
## Add Aetiology group
```{r}
fit_lm3 <- with(
  imp,
  lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles +
       Baseline_endometrium + Aetiology_group)
)
summary(pool(fit_lm3))
```
## Remove BMI and Endo
```{r}
fit_lm4 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group)
)

summary(pool(fit_lm4))
```
## Only AMH and total follicles, no age
```{r}
fit_lm5 <- with(
  imp,
  lm(No_mature_eggs ~ Baseline_AMH + Baseline_total_follicles + Aetiology_group)
)
summary(pool(fit_lm5))
```
## Only AMH and age, no total follicles
```{r}
fit_lm6 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group)
)
summary(pool(fit_lm6))
```
## Only Age and follicles, no AMH
```{r}
fit_lm7 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_total_follicles + Aetiology_group)
)
summary(pool(fit_lm7))
```

## Fit models on one of the MICE datasets and look at evaluation metrics
```{r}
m1_c <- lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles, data = comp1)
m2_c <- lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles + Baseline_endometrium, data = comp1)
m3_c <- lm(No_mature_eggs ~ Age + BMI + Baseline_AMH + Baseline_total_follicles + Baseline_endometrium + Aetiology_group, data = comp1)
m4_c <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group, data = comp1)
m5_c <- lm(No_mature_eggs ~ Baseline_AMH + Baseline_total_follicles + Aetiology_group, data = comp1)
m6_c <- lm(No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group, data = comp1)
m7_c <- lm(No_mature_eggs ~ Age + Baseline_total_follicles + Aetiology_group, data = comp1)

summary(m1_c)$adj.r.squared #0.548
summary(m2_c)$adj.r.squared #0.546 -> adding baseline endometrium provides no further explanation of variance
summary(m3_c)$adj.r.squared #0.622 -> Aetiology group improves explanation of variance
summary(m4_c)$adj.r.squared #0.624 -> removing BMI and endometrium does not affect model much, likely not necessary predictors
summary(m5_c)$adj.r.squared #0.624 -> removing age less of effect
summary(m6_c)$adj.r.squared #0.588 -> don't remove baseline total follicles
summary(m7_c)$adj.r.squared #0.520 -> don't remove baseline AMH

AIC(m1_c, m2_c, m3_c, m4_c, m5_c, m6_c, m7_c) #335.73, 335.74, 329.54, 327.64, 326.82, 331.68, 339.74

shortdata$Aetiology_group <- as.factor((shortdata$Aetiology_group))
m4_c <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group, data = shortdata)
summary(m4_c)


#model 4 is the best - interestly age doesnt add anything extra when there is baseline follicles and baseline AMH, but clinicially relevant and it does not worsen the model so we will keep it in
```

## Interactions

## Age x AMH
```{r}
fit_int1 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Age:Baseline_AMH)
)

summary(pool(fit_int1))
```
## Age x Follicles
```{r}
fit_int2 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Age:Baseline_total_follicles)
)

summary(pool(fit_int2))
```
## AMH x Follicles
```{r}
fit_int3 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Baseline_AMH:Baseline_total_follicles)
)

summary(pool(fit_int3))
```
## AMH x Aetiology group
```{r}
fit_int4 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Baseline_AMH:Aetiology_group)
)

summary(pool(fit_int4))
```
## Age x Aetiology group
```{r}
fit_int5 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Age:Aetiology_group)
)

summary(pool(fit_int5))
```
## Baseline follicles x Aetiology group
```{r}
fit_int6 <- with(
  imp,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Baseline_total_follicles:Aetiology_group)
)

summary(pool(fit_int6))
```

```{r}
m4_int <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Age:Baseline_AMH, data = comp1)
m4_int2 <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Age:Baseline_total_follicles, data = comp1)
m4_int3 <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Baseline_AMH:Baseline_total_follicles, data = comp1)
m4_int4 <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Baseline_AMH:Aetiology_group, data = comp1)
m4_int5 <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Age:Aetiology_group, data = comp1)
m4_int6 <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Baseline_total_follicles:Aetiology_group, data = comp1)

summary(m4_int)$adj.r.squared #0.635
summary(m4_int2)$adj.r.squared #0.617
summary(m4_int3)$adj.r.squared #0.644
summary(m4_int4)$adj.r.squared #0.710
summary(m4_int4)
summary(m4_int5)$adj.r.squared #0.647
summary(m4_int5)
summary(m4_int6)$adj.r.squared #0.660
summary(m4_int6)

AIC(m4_int, m4_int2, m4_int3, m4_int4, m4_int5, m4_int6) #326.97, 329.53, 325.522, 316.25, 326.75, 324.80
```

## Negative Binomial Modelling

```{r}
fit_nb4 <- with(
  imp,
  glm.nb(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group)
)

summary(pool(fit_nb4))

exp(coef(glm.nb(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group, data = comp1)))
```

```{r}
nb4_c <- glm.nb(
  No_mature_eggs ~ Age + Baseline_total_follicles + Baseline_AMH * Aetiology_group,
  data = comp1
)

summary(nb4_c)
AIC(nb4_c) #321.59
```

## actual data
```{r}
nb1 <- glm.nb(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group, data = short_data1)
summary(nb1)

lm1 <- lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles * Aetiology_group, data = short_data1)
summary(lm1)
```

## PROTCOL VARIABLES
```{r}
freshdata <- freshdata %>%
  mutate(
    Stim_regimen = case_when(
      FSH_used == 1 & hMG_used == 0 ~ "FSH_only",
      FSH_used == 0 & hMG_used == 1 ~ "hMG_only",
      FSH_used == 1 & hMG_used == 1 ~ "Dual",
      TRUE ~ "None"
    )
  )

short_data2 <- freshdata %>%
  dplyr::filter(Protocol_type == "Short Antagonist") %>%
  dplyr::select(
    No_mature_eggs,
    Age,
    BMI,
    Baseline_AMH,
    Baseline_total_follicles,
    Baseline_endometrium,
    Aetiology_group,
    GnRH_antagonist_duration,
    GnRH_antagonist_total_dose,
    Stim_regimen,
    FSH_total_dose,
    FSH_initial_dose,
    FSH_dose_changes,
    hMG_total_dose,
    hMG_initial_dose,
    Total_stim_dose,
    hCG_trigger_used,
    GnRH_trigger_used,
    hCG_trigger_dose,
    GnRH_agonist_trigger_dose,
    Last_day_of_stim
  ) %>%
  mutate(
    Aetiology_group = as.factor(Aetiology_group),
    Stim_regimen = as.factor(Stim_regimen)
  )

colSums(is.na(short_data2))
```
# MICE

```{r}
meth <- make.method(short_data2)
pred <- make.predictorMatrix(short_data2)

meth["No_mature_eggs"] <- ""
meth["Age"] <- ""
meth["Baseline_AMH"] <- "pmm"
meth["Baseline_total_follicles"] <- "pmm"
meth["Aetiology_group"] <- "polyreg"
meth["hCG_trigger_dose"] <- "pmm"
meth["GnRH_agonist_trigger_dose"] <- "pmm"

pred["No_mature_eggs", ] <- 0
pred[, "No_mature_eggs"] <- 1

imp2 <- mice(short_data2, m = 20, method = meth, predictorMatrix = pred, seed = 123, printFlag = FALSE)
```

## Expand LM - do stimulation protocols add extra info
```{r}
fit_lm_protocol1 <- with(
  imp2,
  lm(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group +
       Stim_regimen + GnRH_antagonist_total_dose + FSH_initial_dose + hMG_initial_dose + Last_day_of_stim)
)

summary(pool(fit_lm_protocol1))

fit_nb_protocol1 <- with(
  imp2,
  glm.nb(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group +
           Stim_regimen + FSH_initial_dose + hMG_initial_dose)
)

summary(pool(fit_nb_protocol1))
```

```{r}
comp2 <- complete(imp2, 1)

lm_protocol1_c <- lm(
  No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group +
    Stim_regimen + GnRH_antagonist_total_dose + FSH_initial_dose + hMG_initial_dose + Last_day_of_stim,
  data = comp2
)
plot(lm_protocol1_c)
shapiro.test(residuals(lm_protocol1_c)) #0.2807

nb_protocol1_c <- glm.nb(
  No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group + Baseline_total_follicles + Stim_regimen,
  data = comp2
)

summary(nb_protocol1_c) #AIC = 337.86
```


```{r}
lm_protocol2_c <- lm(
  No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group +
    Stim_regimen + FSH_initial_dose + hMG_initial_dose + hCG_trigger_used + GnRH_trigger_used + Aetiology_group:Baseline_AMH,
  data = comp2
)

lm_protocol3_c <- lm(
  No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group +
    Stim_regimen + FSH_initial_dose + hMG_initial_dose + Stim_regimen:Age + Aetiology_group:Baseline_AMH + hCG_trigger_used + GnRH_trigger_used,
  data = comp2
)

lm_protocol4_c <- lm(
  No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles + Aetiology_group + Stim_regimen + FSH_initial_dose + hMG_initial_dose + hCG_trigger_used + GnRH_trigger_used,
  data = comp2
)

summary(lm_protocol2_c)
summary(lm_protocol3_c)
theressummary(lm_protocol4_c)
```

# Check on data
```{r}
summary(lm_protocol1_c) #0.4583
AIC(lm_protocol1_c) #353.889
```


## FSH only patients - from big data, irrespective of short/long antagonist protocol
```{r}
fsh_only <- freshdata %>%
  filter(Stim_regimen == "FSH_only") %>%
  select(
    No_mature_eggs,
    Age,
    BMI,
    Baseline_AMH,
    Baseline_total_follicles,
    Aetiology_group,
    GnRH_antagonist_total_dose,
    GnRH_antagonist_duration,
    FSH_initial_dose,
    FSH_total_dose,
    FSH_dose_changes,
    hCG_trigger_used,
    GnRH_trigger_used,
    Last_day_of_stim
  )

fsh_only <- fsh_only %>%
  mutate(
    Aetiology_group = factor(Aetiology_group),
    hCG_trigger_used = factor(hCG_trigger_used),
    GnRH_trigger_used = factor(GnRH_trigger_used)
  )
```

## Missingness
```{r}
ini <- mice(fsh_only, maxit = 0)
ini$method
ini$predictorMatrix

meth <- ini$method
pred <- ini$predictorMatrix

# Do not impute outcome
meth["No_mature_eggs"] <- ""


meth["BMI"] <- "pmm"
meth["Baseline_AMH"] <- "pmm"
meth["Baseline_total_follicles"] <- "pmm"
meth["Age"] <- ""
meth["Aetiology_group"] <- ""
meth["GnRH_antagonist_total_dose"] <- ""
meth["GnRH_antagonist_duration"] <- ""
meth["FSH_initial_dose"] <- ""
meth["FSH_total_dose"] <- ""
meth["FSH_dose_changes"] <- ""
meth["hCG_trigger_used"] <- ""
meth["GnRH_trigger_used"] <- ""
meth["Last_day_of_stim"] <- ""

imp_fsh <- mice(
  fsh_only,
  m = 20,
  method = meth,
  predictorMatrix = pred,
  seed = 123,
  maxit = 20
)
```

```{r}
fit_fsh1 <- with(
  imp_fsh,
  glm.nb(No_mature_eggs ~ Age + Baseline_AMH + Baseline_total_follicles +
      Aetiology_group + FSH_initial_dose +
      GnRH_antagonist_duration +
      hCG_trigger_used + GnRH_trigger_used + Last_day_of_stim
  )
)

summary(pool(fit_fsh1))
```
## Check on one MICE dataset
```{r}
fsh_comp1 <- complete(imp_fsh, 1)

fsh1 <- glm.nb(No_mature_eggs ~ Age + Aetiology_group + Baseline_total_follicles + FSH_initial_dose, data = fsh_comp1)
summary(fsh1) #AIC  = 319.43

#Model: No_mature_eggs ~ Age + Baseline_AMH * Aetiology_group + Baseline_total_follicles + FSH_initial_dose = AIC = 320
#Interaction Model: No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group + Baseline_total_follicles + FSH_initial_dose = AIC = 320

#removing age and baseline total follicles increases AIC
#when you account for AMH * Aetiology --> FSH has +ve coefficient --> small increase in mature oocyte yield
#when you do not account for AMH * Aetiology --> FSH has -ve coefficient --> small decrease in mature oocyte yield
#AMH is not significant in any variations of this model --> age and total number follicles, and some aetiologies are
#removing AMH decreases AIC --> not massively but is it perhaps not necessary?

#2x Interaction Model: No_mature_eggs ~ Age + Baseline_AMH * Aetiology_group + Baseline_total_follicles * FSH_initial_dose - lowest AIC = 316.92
#Interaction Model: No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group + Baseline_total_follicles * FSH_initial_dose - 2nd lowest AIC. = 318.12
#FSH initial dose has negative coefficient in both models
#total follicles x fsh = significant interaction (p value)

#Model: No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group + Baseline_total_follicles + GnRH_antagonistduration * FSH_initial_dose
#Sig interaction between GnRH antagonist dose x FSH dose
#Sig interaction between GnRH antagonist duration x FSH dose

#Model: No_mature_eggs ~ Age + Baseline_AMH * Aetiology_group + Baseline_total_follicles + GnRH_antagonist_total_dose * FSH_initial_dose
#Sig interaction
#duration x fsh coefficient = larger
```
```{r}
fsh2 <- glm.nb(No_mature_eggs ~ Age + Aetiology_group + Baseline_total_follicles * FSH_initial_dose, data = fsh_comp1)
summary(fsh2) #AIC = 317.06
```

```{r}
fsh3 <- glm.nb(No_mature_eggs ~ Age + Baseline_AMH + Aetiology_group + Baseline_total_follicles * FSH_initial_dose, data = fsh_comp1)
summary(fsh3) #AIC = 318.22
``` 
```{r}
fsh4 <- glm.nb(No_mature_eggs ~ Age + Baseline_AMH * Aetiology_group + Baseline_total_follicles * FSH_initial_dose, data = fsh_comp1)
summary(fsh4) #AIC = 316.92
```
```{r}
fsh_comp1 <- fsh_comp1 %>%
  mutate(
    Age_c = Age - mean(Age, na.rm = TRUE),
    Baseline_total_follicles_c = Baseline_total_follicles - mean(Baseline_total_follicles, na.rm = TRUE),
    FSH_initial_dose_c = FSH_initial_dose - mean(FSH_initial_dose, na.rm = TRUE),
    Baseline_AMH_c = Baseline_AMH - mean(Baseline_AMH, na.rm = TRUE)
  )

fsh2.1 <- glm.nb(No_mature_eggs ~ Age_c + Aetiology_group + Baseline_total_follicles_c * FSH_initial_dose_c, data = fsh_comp1)
summary(fsh2.1) #AIC = 317.06
```

## Checking baseline model on the FSH only data
```{r}
baseline_1 <- lm(No_mature_eggs ~ Baseline_total_follicles * Aetiology_group + Baseline_AMH + Age, data = fsh_comp1)
summary(baseline_1)
```

